### 2020-09-03 paperone_adult_health.Rmd
More changes for analysis and during paper drafting.

---
title: 'Manuscript 1: Adult Health Effects'
author: "Elizabeth Chase"
date: "8/8/2020"
output: pdf_document
---

```{r setup, include=FALSE, echo=TRUE}
library(knitr)
opts_chunk$set(collapse=TRUE, fig.align='center', out.extra = '', fig.pos= "h", tidy=TRUE, tidy.opts=list(blank=TRUE, width.cutoff=40), warning=FALSE,message=FALSE)
options(scipen=999)

rm(list=ls())

library(tidyverse)
library(gridExtra)
library(RColorBrewer)
library(tableone)
library(latex2exp)
library(labelled)
library(kableExtra)
library(tigris)
library(sf)
library(spdep)
library(logistf)

mydata <- readRDS("~/Desktop/Research/Jerel Flint/Analyses/adult_data_clean.rds")
```

We start by selecting our sample:
```{r, echo=TRUE, message=FALSE, warning=FALSE}
mydata$num_time_notice <- case_when(
  mydata$noticed_water_issue=="1 year *AFTER* it was switched" ~ 12,
  mydata$noticed_water_issue=="2 months *AFTER* it was switched" ~ 2,
  mydata$noticed_water_issue=="Unsure/Don't know" ~ NA_real_,
  mydata$noticed_water_issue=="2 years *AFTER* it was switched" ~ 24,
  mydata$noticed_water_issue=="Less than 1 month *AFTER* the water source was switched" ~ 0.5,
  mydata$noticed_water_issue=="1 month *AFTER* it was switched" ~ 1,
  mydata$noticed_water_issue=="3 months *AFTER* it was switched" ~ 3,
  mydata$noticed_water_issue=="1 year and 2 months *AFTER* it was switched" ~ 14,
  mydata$noticed_water_issue=="6 months *AFTER* it was switched" ~ 6,
  mydata$noticed_water_issue=="9 months *AFTER* it was switched" ~ 9,
  mydata$noticed_water_issue=="1 year and 6 months *AFTER* it was switched" ~ 18,
  mydata$noticed_water_issue=="7 months *AFTER* it was switched" ~ 7,
  mydata$noticed_water_issue=="4 years or more *AFTER* it was switched" ~ 4,
  mydata$noticed_water_issue=="10 months *AFTER* it was switched" ~ 10,
  mydata$noticed_water_issue=="5 months *AFTER* it was switched" ~ 5,
  mydata$noticed_water_issue=="11 months *AFTER* it was switched" ~ 11,
  mydata$noticed_water_issue=="4 months *AFTER* it was switched" ~ 4,
  mydata$noticed_water_issue=="1 year and 5 months *AFTER* it was switched" ~ 17,
  mydata$noticed_water_issue=="8 months *AFTER* it was switched" ~ 8,
  mydata$noticed_water_issue=="3 years *AFTER* it was switched" ~ 36,
  mydata$noticed_water_issue=="2 year and 1 month *AFTER* it was switched" ~ 25,
  mydata$noticed_water_issue=="2 year and 6 months *AFTER* it was switched" ~ 30,
  mydata$noticed_water_issue=="1 year and 3 months *AFTER* it was switched" ~ 15,
  mydata$noticed_water_issue=="1 year and 1 month *AFTER* it was switched" ~ 13
)

mydata_complete <- filter(mydata, !is.na(age), !is.na(reth), !is.na(marital_status), !is.na(educ_level), !is.na(public_benefits), !is.na(gender), !is.na(confidants2), !is.na(years_cityflint), !is.na(current_zip), !is.na(blood_test_after_april_2014), !is.na(blood_test_before_april_2014), !is.na(elevated_lead_before_april_2014), !is.na(elevated_lead_after_april_2014), !is.na(rashes_before_april_2014), !is.na(rashes_after_april_2014), !is.na(hairloss_before_april_2014), !is.na(hairloss_after_april_2014), !is.na(nausea_before_april_2014), !is.na(nausea_after_april_2014), !is.na(irritable_before_april_2014), !is.na(irritable_after_april_2014), !is.na(ptsdscore), !is.na(phq), gender != "Transgender Woman", !is.na(num_time_notice))

mydata_complete$collapsemarital <- case_when(
  mydata_complete$marital_status=="Single (never married)" ~ "Single (never married)",
  mydata_complete$marital_status=="Married" ~ "Married", 
  mydata_complete$marital_status=="Separated" | mydata_complete$marital_status=="Divorced" | mydata_complete$marital_status=="Widowed" ~ "Separated"
)

mydata_complete$overall_collapse <- rowSums(select(mydata_complete, rash_collapse, hair_collapse, irritable_collapse, nausea_collapse))

mydata_complete$bloodtest_collapse <- ifelse(mydata_complete$blood_test_after_april_2014=="Yes", 1, 0)

mydata_complete$ptsd_bin <- as.numeric(mydata_complete$ptsdscore >= 3)
mydata_complete$phq_bin <- as.numeric(mydata_complete$phq >= 6)

mydata_complete$ptsd_bin <- factor(mydata_complete$ptsd_bin, level=c(0,1), labels=c("No", "Yes"))
mydata_complete$phq_bin <- factor(mydata_complete$phq_bin, level=c(0,1), labels=c("No", "Yes"))
```

We assemble some descriptives and maps on the complete data:
```{r, echo=TRUE, message=FALSE, warning=FALSE}
zcta1 <- zctas(cb = TRUE, starts_with = c("4850", "48532"), class = "sf") %>% filter(ZCTA5CE10 != "48509")
flintdat <- data.frame("zip" = zcta1$ZCTA5CE10, "geometry" = zcta1$geometry)

fils <- unzip("~/Desktop/Research/Jerel Flint/Analyses/tl_2019_26_place.zip")
flint_boundary <- st_read(fils[4]) %>% filter(NAME=="Flint")

totalpop <- data.frame(table(mydata_complete$current_zip))
colnames(totalpop) <- c("current_zip", "pop")

flintdat2 <- full_join(flintdat, totalpop, by=c("zip" = "current_zip"))

popmap <- ggplot(data=flintdat2) + geom_sf(data=flintdat2, aes(geometry = geometry, fill = factor(pop)), size=0, alpha = 0.6) + geom_sf(data=flint_boundary, aes(geometry=geometry), fill=NA, color = "black") + geom_sf_label(aes(geometry = geometry, label = paste0(zip, " (", pop, ")"))) + xlab(" ") + ylab(" ") + theme(rect = element_blank(), axis.ticks = element_blank(), axis.line=element_blank(),axis.text.x=element_blank(), axis.text.y=element_blank(), legend.position = "none") + scale_fill_grey()

popmap

ggsave(filename = "mymap.pdf", plot = popmap, width = 8, height = 8)

var_label(mydata_complete$collapseeduc) <- "Education level"
var_label(mydata_complete$public_benefits) <- "Receives public benefits"
var_label(mydata_complete$collapsemarital) <- "Marital status"
var_label(mydata_complete$num_time_notice) <- "How long after water switch to notice water issue? (months)"
var_label(mydata_complete$blood_test_after_april_2014) <- "Received blood lead test"
var_label(mydata_complete$elevated_lead_after_april_2014) <- "Had elevated lead"
var_label(mydata_complete$rashes_after_april_2014) <- "Had rashes"
var_label(mydata_complete$hairloss_after_april_2014) <- "Had hair loss"
var_label(mydata_complete$nausea_after_april_2014) <- "Had nausea"
var_label(mydata_complete$irritable_after_april_2014) <- "Was irritable"
var_label(mydata_complete$ptsd_bin) <- "PTSD"
var_label(mydata_complete$phq_bin) <- "Depression/Anxiety"

table_predictors <- CreateTableOne(vars=c("age", "gender", "collapseeduc", "public_benefits", "collapsemarital", "number_children", "years_cityflint", "confidants2"), strata = "collapserace", data=mydata_complete, argsNormal = list(NULL), argsNonNormal = list(var.equal = TRUE))

table_outcomes <- CreateTableOne(vars=c("num_time_notice", "blood_test_after_april_2014", "elevated_lead_after_april_2014", "rashes_after_april_2014", "hairloss_after_april_2014", "nausea_after_april_2014", "irritable_after_april_2014", "overall_collapse", "ptsd_bin", "phq_bin"), strata = "collapserace", data=mydata_complete, argsNormal = list(NULL), argsNonNormal = list(var.equal = TRUE))

kable(print(table_predictors, contDigits=1, varLabel = T, printToggle=FALSE), booktabs=T) %>% add_indent(c(5:9, 11:13, 15:17)) %>% kable_styling(position="center")

kable(print(table_outcomes, contDigits=1, varLabel = T, printToggle=FALSE), booktabs=T) %>% add_indent(c(4:6, 8:10, 12:14, 16:18, 20:22, 24:26)) %>% kable_styling(position="center")

table_predictors_notstrat <- CreateTableOne(vars=c("age", "gender", "collapseeduc", "public_benefits", "collapsemarital", "number_children", "years_cityflint", "confidants2"), data=mydata_complete, argsNormal = list(NULL), argsNonNormal = list(var.equal = TRUE))

table_outcomes_notstrat <- CreateTableOne(vars=c("num_time_notice", "blood_test_after_april_2014", "elevated_lead_after_april_2014", "rashes_after_april_2014", "hairloss_after_april_2014", "nausea_after_april_2014", "irritable_after_april_2014", "overall_collapse", "ptsd_bin", "phq_bin"), data=mydata_complete, argsNormal = list(NULL), argsNonNormal = list(var.equal = TRUE))

kable(print(table_predictors_notstrat, contDigits=1, varLabel = T, printToggle=FALSE), booktabs=T) %>% add_indent(c(5:9, 11:13, 15:17)) %>% kable_styling(position="center")

kable(print(table_outcomes_notstrat, contDigits=1, varLabel = T, printToggle=FALSE), booktabs=T) %>% add_indent(c(4:6, 8:10, 12:14, 16:18, 20:22, 24:26)) %>% kable_styling(position="center")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
knitr::include_graphics("~/Desktop/Research/Jerel Flint/Analyses/sample_inclusion.png")
```

Now, we conduct Moran's I tests grouped by zipcode to reassure ourselves that we can just adjust for zipcode and not do a full-on spatial analysis:
```{r, echo=FALSE, message=FALSE, warning=FALSE}
zip_symp <- mydata_complete %>% dplyr::group_by(current_zip) %>% dplyr::summarize(elev_blood = sum(elevated_collapse, na.rm=TRUE), rash = sum(rash_collapse, na.rm=TRUE), nausea = sum(nausea_collapse, na.rm=TRUE), hairloss = sum(hair_collapse, na.rm=TRUE), irritable = sum(irritable_collapse, na.rm=TRUE), overall = sum(overall_collapse, na.rm=TRUE), ptsd = sum(ptsd_bin=="Yes", na.rm=TRUE), phq = sum(phq_bin=="Yes", na.rm=TRUE), get_test = sum(bloodtest_collapse, na.rm=TRUE), notice_issue = mean(num_time_notice, na.rm=TRUE), black = sum(collapserace=="Black", na.rm=TRUE), other = sum(collapserace=="Other", na.rm=TRUE),  l9 = sum(collapseeduc=="Less than high school", na.rm=TRUE), hs = sum(collapseeduc=="High school", na.rm=TRUE), assc = sum(collapseeduc=="Associate's", na.rm=TRUE), somecoll = sum(collapseeduc=="Some college", na.rm=TRUE), female = sum(gender!="Male", na.rm=TRUE), age = mean(age, na.rm=TRUE), pubben = sum(public_benefits=="Yes", na.rm=TRUE), pubbendk = sum(public_benefits=="Don't know/Unsure", na.rm=TRUE), network = mean(confidants2, na.rm=TRUE))

flintdat3 <- full_join(flintdat2, zip_symp, by=c("zip" = "current_zip"))

flintdat3$l9 <- flintdat3$l9/flintdat3$pop
flintdat3$hs <- flintdat3$hs/flintdat3$pop
flintdat3$assc <- flintdat3$assc/flintdat3$pop
flintdat3$somecoll <- flintdat3$somecoll/flintdat3$pop
flintdat3$black <- flintdat3$black/flintdat3$pop
flintdat3$other <- flintdat3$other/flintdat3$pop
flintdat3$female <- flintdat3$female/flintdat3$pop
flintdat3$pubben <- flintdat3$pubben/flintdat3$pop
flintdat3$pubbendk <- flintdat3$pubbendk/flintdat3$pop

flint.nb <- poly2nb(flintdat3$geometry)
flint.list.w <- nb2listw(flint.nb)

elevblood <- moran.test(flintdat3$elev_blood/flintdat3$pop, listw=flint.list.w, randomisation=FALSE)
elevblood
elevblood_adj <- lm.morantest(lm(flintdat3$elev_blood/flintdat3$pop~ flintdat3$black + flintdat3$other + flintdat3$female + flintdat3$age),listw=flint.list.w)
elevblood_adj

rash <- moran.test(flintdat3$rash/flintdat3$pop, listw=flint.list.w, randomisation=FALSE)
rash
rash_adj <- lm.morantest(lm(flintdat3$rash/flintdat3$pop ~ flintdat3$black + flintdat3$other + flintdat3$female + flintdat3$age),listw=flint.list.w)
rash_adj

nausea <- moran.test(flintdat3$nausea/flintdat3$pop, listw=flint.list.w, randomisation=FALSE)
nausea
nausea_adj <- lm.morantest(lm(flintdat3$nausea/flintdat3$pop ~ flintdat3$black + flintdat3$other + flintdat3$female + flintdat3$age),listw=flint.list.w)
nausea_adj

hairloss <- moran.test(flintdat3$hairloss/flintdat3$pop, listw=flint.list.w, randomisation=FALSE)
hairloss
hairloss_adj <- lm.morantest(lm(flintdat3$hairloss/flintdat3$pop ~ flintdat3$black + flintdat3$other + flintdat3$female + flintdat3$age),listw=flint.list.w)
hairloss_adj

irritable <- moran.test(flintdat3$irritable/flintdat3$pop, listw=flint.list.w, randomisation=FALSE)
irritable
irritable_adj <- lm.morantest(lm(flintdat3$irritable/flintdat3$pop ~ flintdat3$black + flintdat3$other + flintdat3$female + flintdat3$age),listw=flint.list.w)
irritable_adj

overall <- moran.test(flintdat3$overall/flintdat3$pop, listw=flint.list.w, randomisation=FALSE)
overall
overall_adj <- lm.morantest(lm(flintdat3$overall/flintdat3$pop ~ flintdat3$black + flintdat3$other + flintdat3$female + flintdat3$age),listw=flint.list.w)
overall_adj

ptsd <- moran.test(flintdat3$ptsd/flintdat3$pop, listw=flint.list.w, randomisation=FALSE)
ptsd
ptsd_adj <- lm.morantest(lm(flintdat3$ptsd/flintdat3$pop ~ flintdat3$black + flintdat3$other + flintdat3$female + flintdat3$age),listw=flint.list.w)
ptsd_adj

phq <- moran.test(flintdat3$phq/flintdat3$pop, listw=flint.list.w, randomisation=FALSE)
phq
phq_adj <- lm.morantest(lm(flintdat3$phq/flintdat3$pop ~ flintdat3$black + flintdat3$other + flintdat3$female + flintdat3$age),listw=flint.list.w)
phq_adj

get_tested <- moran.test(flintdat3$get_test/flintdat3$pop, listw=flint.list.w, randomisation=FALSE)
get_tested
get_tested_adj <- lm.morantest(lm(flintdat3$get_test/flintdat3$pop ~ flintdat3$black + flintdat3$other + flintdat3$female + flintdat3$age),listw=flint.list.w)
get_tested_adj

notice <- moran.test(flintdat3$notice_issue, listw=flint.list.w, randomisation=FALSE)
notice
notice_adj <- lm.morantest(lm(flintdat3$notice_issue ~ flintdat3$black + flintdat3$other + flintdat3$female + flintdat3$age),listw=flint.list.w)
notice_adj
```

There is some evidence of spatial correlation for nausea and hairloss, but once we adjust for baseline covariates it mostly goes away. Given the constraints of this analysis, we're going to just adjust for zipcode in our models and call it a day.

We fit our models according to the causal diagram given here:
```{r, echo=FALSE, message=FALSE, warning=FALSE}
knitr::include_graphics("~/Desktop/Research/Jerel Flint/Analyses/mediation.png")
```

As we know, when trying to understand the effect of a mediated effect on some outcome, we DO NOT adjust for the mediator. However, when trying to understand the effect of the mediator on the outcome, we DO adjust for the mediated effect. Therefore, we have to fit several models for our different outcomes, adjusting for different things depending on what the predictor of interest is. All told, we consider 10 different outcomes (PTSD, PHQ, rashes, hair loss, nausea, irritability, overall number of symptoms, elevated blood lead, getting tested, time to noticing issues) and consider the classes of predictors: demographics/baseline, time to noticing issues/getting tested, elevated blood lead, and emergence of symptoms. This results in us fitting 17 different models, given here:
```{r, echo=FALSE, message=FALSE, warning=FALSE}
mydata_complete$collapsezip <- ifelse(mydata_complete$current_zip==48502, 2, mydata_complete$current_zip)
mydata_complete$collapsezip <- factor(mydata_complete$collapsezip, levels=c(2:7), labels=c("48502/48503", "48504", "48505", "48506", "48507", "48532"))
mydata_complete$collapsepubben <- ifelse(mydata_complete$public_benefits=="Yes", "Yes", "No")

ptsd_dem <- glm(ptsd_bin ~ age + gender + collapserace + collapsepubben + number_children + years_cityflint + confidants2 + collapsezip, data=mydata_complete, family="binomial")

summary(ptsd_dem)

phq_dem <- glm(phq_bin ~ age + gender + collapserace + collapsepubben + number_children + years_cityflint + confidants2 + collapsezip, data=mydata_complete, family="binomial")

summary(phq_dem)

rash_dem <- glm(rash_collapse ~ age + gender + collapserace + collapsepubben + number_children + years_cityflint + confidants2 + collapsezip, data=mydata_complete, family="binomial")

summary(rash_dem)

hairloss_dem <- glm(hair_collapse ~ age + gender + collapserace + collapsepubben + number_children + years_cityflint + confidants2 + collapsezip, data=mydata_complete, family="binomial")

summary(hairloss_dem)

nausea_dem <- glm(nausea_collapse ~ age + gender + collapserace + collapsepubben + number_children + years_cityflint + confidants2 + collapsezip, data=mydata_complete, family="binomial")

summary(nausea_dem)

irritable_dem <- glm(irritable_collapse ~ age + gender + collapserace + collapsepubben + number_children + years_cityflint + confidants2 + collapsezip, data=mydata_complete, family="binomial")

summary(irritable_dem)

overall_dem <- glm(overall_collapse ~ age + gender + collapserace + collapsepubben + number_children + years_cityflint + confidants2 + collapsezip, data=mydata_complete, family="poisson")

summary(overall_dem)

bloodlead_dem <- logistf(elevated_collapse ~ age + gender + collapserace + collapsepubben + number_children + years_cityflint + confidants2 + collapsezip, data = mydata_complete)

test1 <- summary(bloodlead_dem)

gettest_dem <- glm(bloodtest_collapse ~ age + gender + collapserace + collapsepubben + number_children + years_cityflint + confidants2 + collapsezip, data=mydata_complete, family="binomial")

summary(gettest_dem)

time_dem <- lm(log(num_time_notice) ~ age + gender + collapserace + collapsepubben + number_children + years_cityflint + confidants2 + collapsezip, data=mydata_complete)

summary(time_dem)

mydata_complete$rash_collapse_before <- ifelse(mydata_complete$rashes_before_april_2014=="Yes", "Yes", "No")
mydata_complete$hair_collapse_before <- ifelse(mydata_complete$hairloss_before_april_2014=="Yes", "Yes", "No")
mydata_complete$nausea_collapse_before <- ifelse(mydata_complete$nausea_before_april_2014=="Yes", "Yes", "No")
mydata_complete$irritable_collapse_before <- ifelse(mydata_complete$irritable_before_april_2014=="Yes", "Yes", "No")
mydata_complete$bloodtest_collapse_before <- ifelse(mydata_complete$blood_test_before_april_2014=="Yes", "Yes", "No")

ptsd_symp <- glm(ptsd_bin ~ age + gender + collapserace + collapsepubben + number_children + years_cityflint + confidants2 + collapsezip + overall_collapse + elevated_collapse + bloodtest_collapse + num_time_notice, data=mydata_complete, family="binomial")

summary(ptsd_symp)

phq_symp <- glm(phq_bin ~ age + gender + collapserace + collapsepubben + number_children + years_cityflint + confidants2 + collapsezip + overall_collapse + elevated_collapse + bloodtest_collapse + num_time_notice, data=mydata_complete, family="binomial")

summary(phq_symp)

rash_symp <- logistf(rash_collapse ~ age + gender + collapserace + collapsepubben + number_children + years_cityflint + confidants2 + collapsezip + hair_collapse + nausea_collapse + irritable_collapse + elevated_collapse + bloodtest_collapse + num_time_notice + rash_collapse_before, data=mydata_complete)

test2 <- summary(rash_symp)

hairloss_symp <- logistf(hair_collapse ~ age + gender + collapserace + collapsepubben + number_children + years_cityflint + confidants2 + collapsezip + rash_collapse + nausea_collapse + irritable_collapse + elevated_collapse + bloodtest_collapse + num_time_notice + hair_collapse_before, data=mydata_complete)

test3 <- summary(hairloss_symp)

nausea_symp <- logistf(nausea_collapse ~ age + gender + collapserace + collapsepubben + number_children + years_cityflint + confidants2 + collapsezip + rash_collapse + hair_collapse + irritable_collapse + elevated_collapse + bloodtest_collapse + num_time_notice + nausea_collapse_before, data=mydata_complete)

test4 <- summary(nausea_symp)

irritable_symp <- logistf(irritable_collapse ~ age + gender + collapserace + collapsepubben + number_children + years_cityflint + confidants2 + collapsezip + rash_collapse + hair_collapse + nausea_collapse + elevated_collapse + bloodtest_collapse + num_time_notice + irritable_collapse_before, data=mydata_complete)

test5 <- summary(irritable_symp)

overall_bloodlead <- glm(overall_collapse ~ age + gender + collapserace + collapsepubben + number_children + years_cityflint + confidants2 + collapsezip + elevated_collapse + bloodtest_collapse + num_time_notice + rash_collapse_before + hair_collapse_before + nausea_collapse_before + irritable_collapse_before, data=mydata_complete, family="poisson")

summary(overall_bloodlead)
```

Since we fit quite a few models, we use the FDR to adjust for our multiple testing.
```{r, echo=FALSE, message=FALSE, warning=FALSE}
#Assembling a data frame of desired coefficients, confidence intervals, and p-values:
myresults <- data.frame("Outcome" = c(rep("PTSD", length(summary(ptsd_dem)$coefficients[-1,1])), rep("PHQ", length(summary(phq_dem)$coefficients[-1,1])), rep("Rashes", length(summary(rash_dem)$coefficients[-1,1])), rep("Hairloss", length(summary(hairloss_dem)$coefficients[-1,1])), rep("Nausea", length(summary(nausea_dem)$coefficients[-1,1])), rep("Irritable", length(summary(irritable_dem)$coefficients[-1,1])), rep("Overall", length(summary(overall_dem)$coefficients[-1,1])), rep("Elevated lead", length(test1$coefficients[-1])), rep("Get tested", length(summary(gettest_dem)$coefficients[-1,1])), rep("Time to notice", length(summary(time_dem)$coefficients[-1,1])), rep("PTSD", length(summary(ptsd_symp)$coefficients[15:18,1])), rep("PHQ", length(summary(phq_symp)$coefficients[15:18,1])), rep("Rashes", length(test2$coefficients[15:20])), rep("Hairloss", length(test3$coefficients[15:20])), rep("Nausea", length(test4$coefficients[15:20])), rep("Irritable", length(test5$coefficients[15:20])), rep("Overall", length(summary(overall_bloodlead)$coefficients[15:17,1]))))

myresults$Predictor <- c(rownames(summary(ptsd_dem)$coefficients[-1,]), rownames(summary(phq_dem)$coefficients[-1,]), rownames(summary(rash_dem)$coefficients[-1,]), rownames(summary(hairloss_dem)$coefficients[-1,]), rownames(summary(nausea_dem)$coefficients[-1,]), rownames(summary(irritable_dem)$coefficients[-1,]), rownames(summary(overall_dem)$coefficients[-1,]), names(test1$coefficients[-1]), rownames(summary(gettest_dem)$coefficients[-1,]), rownames(summary(time_dem)$coefficients[-1,]), rownames(summary(ptsd_symp)$coefficients[15:18,]), rownames(summary(phq_symp)$coefficients[15:18,]), names(test2$coefficients[15:20]), names(test3$coefficients[15:20]), names(test4$coefficients[15:20]), names(test5$coefficients[15:20]), rownames(summary(overall_bloodlead)$coefficients[15:17,]))

myresults$Coefficient <- c(summary(ptsd_dem)$coefficients[-1,1], summary(phq_dem)$coefficients[-1,1], summary(rash_dem)$coefficients[-1,1], summary(hairloss_dem)$coefficients[-1,1], summary(nausea_dem)$coefficients[-1,1], summary(irritable_dem)$coefficients[-1,1], summary(overall_dem)$coefficients[-1,1], test1$coefficients[-1], summary(gettest_dem)$coefficients[-1,1], summary(time_dem)$coefficients[-1,1], summary(ptsd_symp)$coefficients[15:18,1], summary(phq_symp)$coefficients[15:18,1], test2$coefficients[15:20], test3$coefficients[15:20], test4$coefficients[15:20], test5$coefficients[15:20], summary(overall_bloodlead)$coefficients[15:17,1])

myresults$Lower <- c(summary(ptsd_dem)$coefficients[-1,1]-1.96*summary(ptsd_dem)$coefficients[-1,2], summary(phq_dem)$coefficients[-1,1]-1.96*summary(phq_dem)$coefficients[-1,2], summary(rash_dem)$coefficients[-1,1]-1.96*summary(rash_dem)$coefficients[-1,2], summary(hairloss_dem)$coefficients[-1,1]-1.96*summary(hairloss_dem)$coefficients[-1,2], summary(nausea_dem)$coefficients[-1,1]-1.96*summary(nausea_dem)$coefficients[-1,2], summary(irritable_dem)$coefficients[-1,1]-1.96*summary(irritable_dem)$coefficients[-1,2], summary(overall_dem)$coefficients[-1,1]-1.96*summary(overall_dem)$coefficients[-1,2], test1$ci.lower[-1], summary(gettest_dem)$coefficients[-1,1]-1.96*summary(gettest_dem)$coefficients[-1,2], summary(time_dem)$coefficients[-1,1]-1.96*summary(time_dem)$coefficients[-1,2], summary(ptsd_symp)$coefficients[15:18,1]-1.96*summary(ptsd_symp)$coefficients[15:18,2], summary(phq_symp)$coefficients[15:18,1]-1.96*summary(phq_symp)$coefficients[15:18,2], test2$ci.lower[15:20], test3$ci.lower[15:20], test4$ci.lower[15:20], test5$ci.lower[15:20], summary(overall_bloodlead)$coefficients[15:17,1]-1.96*summary(overall_bloodlead)$coefficients[15:17,2])

myresults$Upper <- c(summary(ptsd_dem)$coefficients[-1,1]+1.96*summary(ptsd_dem)$coefficients[-1,2], summary(phq_dem)$coefficients[-1,1]+1.96*summary(phq_dem)$coefficients[-1,2], summary(rash_dem)$coefficients[-1,1]+1.96*summary(rash_dem)$coefficients[-1,2], summary(hairloss_dem)$coefficients[-1,1]+1.96*summary(hairloss_dem)$coefficients[-1,2], summary(nausea_dem)$coefficients[-1,1]+1.96*summary(nausea_dem)$coefficients[-1,2], summary(irritable_dem)$coefficients[-1,1]+1.96*summary(irritable_dem)$coefficients[-1,2], summary(overall_dem)$coefficients[-1,1]+1.96*summary(overall_dem)$coefficients[-1,2], test1$ci.upper[-1], summary(gettest_dem)$coefficients[-1,1]+1.96*summary(gettest_dem)$coefficients[-1,2], summary(time_dem)$coefficients[-1,1]+1.96*summary(time_dem)$coefficients[-1,2], summary(ptsd_symp)$coefficients[15:18,1]+1.96*summary(ptsd_symp)$coefficients[15:18,2], summary(phq_symp)$coefficients[15:18,1]+1.96*summary(phq_symp)$coefficients[15:18,2], test2$ci.upper[15:20], test3$ci.upper[15:20], test4$ci.upper[15:20], test5$ci.upper[15:20], summary(overall_bloodlead)$coefficients[15:17,1]+1.96*summary(overall_bloodlead)$coefficients[15:17,2])

myresults$pvalue <- c(summary(ptsd_dem)$coefficients[-1,4], summary(phq_dem)$coefficients[-1,4], summary(rash_dem)$coefficients[-1,4], summary(hairloss_dem)$coefficients[-1,4], summary(nausea_dem)$coefficients[-1,4], summary(irritable_dem)$coefficients[-1,4], summary(overall_dem)$coefficients[-1,4], test1$prob[-1], summary(gettest_dem)$coefficients[-1,4], summary(time_dem)$coefficients[-1,4], summary(ptsd_symp)$coefficients[15:18,4], summary(phq_symp)$coefficients[15:18,4], test2$prob[15:20], test3$prob[15:20], test4$prob[15:20], test5$prob[15:20], summary(overall_bloodlead)$coefficients[15:17,4])

myresults$BHpvalue <- p.adjust(myresults$pvalue, method="BH")
myresults$sig <- ifelse(myresults$BHpvalue <=0.05, "LOOK", NA_character_)

myresults$Predictor_clean <- case_when(
  myresults$Predictor=="age" ~ "Age",
  myresults$Predictor=="genderMale" ~ "Male",
  myresults$Predictor=="collapseraceOther" ~ "Other race",
  myresults$Predictor=="collapseraceWhite" ~ "White race",
  myresults$Predictor=="collapseeducHigh school" ~ "High school graduate",
  myresults$Predictor=="collapseeducSome college" ~ "Some college",
  myresults$Predictor=="collapseeducAssociate's" ~ "Associate's degree",
  myresults$Predictor=="collapseeducCollege degree or more" ~ "College degree or more",
  myresults$Predictor=="collapsepubbenYes" ~ "Receives public benefits",
  myresults$Predictor=="collapsemaritalSeparated" ~ "Separated",
  myresults$Predictor=="collapsemaritalSingle (never married)" ~ "Single",
  myresults$Predictor=="years_cityflint" ~ "Years in Flint",
  myresults$Predictor=="confidants2" ~ "Network size",
  myresults$Predictor=="collapsezip48504" ~ "Zipcode 48504",
  myresults$Predictor=="collapsezip48505" ~ "Zipcode 48505",
  myresults$Predictor=="collapsezip48506" ~ "Zipcode 48506",
  myresults$Predictor=="collapsezip48507" ~ "Zipcode 48507",
  myresults$Predictor=="collapsezip48532" ~ "Zipcode 48532",
  myresults$Predictor=="rash_collapse" ~ "Rashes",
  myresults$Predictor=="hair_collapse" ~ "Hair loss",
  myresults$Predictor=="nausea_collapse" ~ "Nausea",
  myresults$Predictor=="irritable_collapse" ~ "Irritability",
  myresults$Predictor=="elevated_collapse" ~ "Elevated BLL",
  myresults$Predictor=="bloodtest_collapse" ~ "Screened for lead",
  myresults$Predictor=="num_time_notice" ~ "Months to notice problem"
)

myresults$Predictor_clean <- factor(myresults$Predictor_clean, levels= c("Age", "Male", "Other race", "White race", "High school graduate", "Some college", "Associate's degree", "College degree or more", "Receives public benefits", "Separated", "Single", "Years in Flint", "Network size", "Zipcode 48504", "Zipcode 48505", "Zipcode 48506", "Zipcode 48507", "Zipcode 48532", "Months to notice problem", "Screened for lead", "Elevated BLL", "Rashes", "Hair loss", "Nausea", "Irritability"), ordered=TRUE)

table3_results <- filter(myresults, Outcome=="Time to notice" | Outcome=="Get tested" | Outcome=="Elevated lead") %>%
  mutate(Label = paste(round(exp(Coefficient), digits=1), " (", round(exp(Lower), digits=1), ",", round(exp(Upper), digits=1), ")", sep=""), Pvalue = round(BHpvalue, digits=2)) %>% select(Outcome, Predictor_clean, Label, Pvalue)

table4_results <- filter(myresults, Outcome=="Rashes" | Outcome=="Hairloss" | Outcome=="Nausea" | Outcome=="Irritable") %>%
  mutate(Label = paste(round(exp(Coefficient), digits=1), " (", round(exp(Lower), digits=1), ",", round(exp(Upper), digits=1), ")", sep=""), Pvalue = round(BHpvalue, digits=2)) %>% select(Outcome, Predictor_clean, Label, Pvalue)

table4_resultsb <- filter(myresults, Outcome=="Overall") %>%
  mutate(Label = paste(round(Coefficient, digits=1), " (", round(Lower, digits=1), ",", round(Upper, digits=1), ")", sep=""), Pvalue = round(BHpvalue, digits=2)) %>% select(Outcome, Predictor_clean, Label, Pvalue)

table5_results <- filter(myresults, Outcome=="PTSD" | Outcome=="PHQ") %>%
  mutate(Label = paste(round(exp(Coefficient), digits=1), " (", round(exp(Lower), digits=1), ",", round(exp(Upper), digits=1), ")", sep=""), Pvalue = round(BHpvalue, digits=2)) %>% select(Outcome, Predictor_clean, Label, Pvalue)

#ptsd_plot <- ggplot(myresults[myresults$Outcome=="PTSD",], aes(fct_rev(Predictor_clean), Coefficient)) + geom_point() + geom_errorbar(aes(ymin=Lower, ymax=Upper, width=0.1)) + coord_flip(ylim=c(-2.35, 5.5)) + labs(y = "Odds Ratio and 95% CI", x = "", caption=TeX('Protective $\\leftarrow \\rightarrow$ Harmful' )) + geom_hline(yintercept=0, linetype="dashed", color = "black") + theme(panel.background = element_rect(fill="white"), panel.grid.major = element_blank(), panel.grid.minor=element_blank(), legend.position="bottom", text=element_text(size=14), plot.caption = element_text(size=14, hjust = 0.15)) + geom_rect(fill="white", ymin=4.7, ymax=Inf, xmin=-Inf, xmax=Inf) + annotate(geom = 'text', label = myresults$Label[myresults$Outcome=="PTSD"], x = myresults$Predictor_clean[myresults$Outcome=="PTSD"], y = 3.5, hjust = 0, vjust = 0.5) + scale_y_continuous(breaks=log(c(0.1, 0.25, 0.5, 1, 2, 4, 10)), labels=c("0.1","0.25", "0.5", "1.0", "2.0", "4.0", "10.0")) + ggtitle("Outcome: Trauma Symptoms")

#ptsd_plot

#phq_plot <- ggplot(myresults[myresults$Outcome=="PHQ",], aes(fct_rev(Predictor_clean), Coefficient)) + geom_point() + geom_errorbar(aes(ymin=Lower, ymax=Upper, width=0.1)) + coord_flip(ylim=c(-2.35, 5.5)) + labs(y = "Odds Ratio and 95% CI", x = "", caption=TeX('Protective $\\leftarrow \\rightarrow$ Harmful' )) + geom_hline(yintercept=0, linetype="dashed", color = "black") + theme(panel.background = element_rect(fill="white"), panel.grid.major = element_blank(), panel.grid.minor=element_blank(), legend.position="bottom", text=element_text(size=14), plot.caption = element_text(size=14, hjust = 0.15), axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank()) + geom_rect(fill="white", ymin=4.7, ymax=Inf, xmin=-Inf, xmax=Inf) + annotate(geom = 'text', label = myresults$Label[myresults$Outcome=="PHQ"], x = myresults$Predictor_clean[myresults$Outcome=="PHQ"], y = 3.5, hjust = 0, vjust = 0.5) + scale_y_continuous(breaks=log(c(0.1, 0.25, 0.5, 1, 2, 4, 10)), labels=c("0.1","0.25", "0.5", "1.0", "2.0", "4.0", "10.0")) + ggtitle("Outcome: Depression/Anxiety")

#phq_plot

#mental_plots <- grid.arrange(ptsd_plot, phq_plot, nrow=1, widths = c(6, 3))

```