### 2021-02-02 paperone_adult_health.Rmd
Changes to meet reviewer comments for JUH.

---
title: 'Manuscript 1: Adult Health Effects'
author: "Elizabeth Chase"
date: "8/8/2020"
output: pdf_document
---

```{r setup, include=FALSE, echo=TRUE}
library(knitr)
opts_chunk$set(collapse=TRUE, fig.align='center', out.extra = '', fig.pos= "h", tidy=TRUE, tidy.opts=list(blank=TRUE, width.cutoff=40), warning=FALSE,message=FALSE)
options(scipen=999)

rm(list=ls())

library(tidyverse)
library(gridExtra)
library(RColorBrewer)
library(tableone)
library(latex2exp)
library(labelled)
library(kableExtra)
library(tigris)
library(sf)
library(spdep)
library(logistf)
library(mice)

mydata <- readRDS("~/Desktop/Research/Jerel Flint/Analyses/adult_data_clean.rds")
```

We start by selecting our sample:
```{r, echo=TRUE, message=FALSE, warning=FALSE}
mydata$num_time_notice <- case_when(
  mydata$noticed_water_issue=="1 year *AFTER* it was switched" ~ 12,
  mydata$noticed_water_issue=="2 months *AFTER* it was switched" ~ 2,
  mydata$noticed_water_issue=="Unsure/Don't know" ~ NA_real_,
  mydata$noticed_water_issue=="2 years *AFTER* it was switched" ~ 24,
  mydata$noticed_water_issue=="Less than 1 month *AFTER* the water source was switched" ~ 0.5,
  mydata$noticed_water_issue=="1 month *AFTER* it was switched" ~ 1,
  mydata$noticed_water_issue=="3 months *AFTER* it was switched" ~ 3,
  mydata$noticed_water_issue=="1 year and 2 months *AFTER* it was switched" ~ 14,
  mydata$noticed_water_issue=="6 months *AFTER* it was switched" ~ 6,
  mydata$noticed_water_issue=="9 months *AFTER* it was switched" ~ 9,
  mydata$noticed_water_issue=="1 year and 6 months *AFTER* it was switched" ~ 18,
  mydata$noticed_water_issue=="7 months *AFTER* it was switched" ~ 7,
  mydata$noticed_water_issue=="4 years or more *AFTER* it was switched" ~ 4,
  mydata$noticed_water_issue=="10 months *AFTER* it was switched" ~ 10,
  mydata$noticed_water_issue=="5 months *AFTER* it was switched" ~ 5,
  mydata$noticed_water_issue=="11 months *AFTER* it was switched" ~ 11,
  mydata$noticed_water_issue=="4 months *AFTER* it was switched" ~ 4,
  mydata$noticed_water_issue=="1 year and 5 months *AFTER* it was switched" ~ 17,
  mydata$noticed_water_issue=="8 months *AFTER* it was switched" ~ 8,
  mydata$noticed_water_issue=="3 years *AFTER* it was switched" ~ 36,
  mydata$noticed_water_issue=="2 year and 1 month *AFTER* it was switched" ~ 25,
  mydata$noticed_water_issue=="2 year and 6 months *AFTER* it was switched" ~ 30,
  mydata$noticed_water_issue=="1 year and 3 months *AFTER* it was switched" ~ 15,
  mydata$noticed_water_issue=="1 year and 1 month *AFTER* it was switched" ~ 13
)

mydata$collapsemarital <- case_when(
  mydata$marital_status=="Single (never married)" ~ "Single (never married)",
  mydata$marital_status=="Married" ~ "Married", 
  mydata$marital_status=="Separated" | mydata$marital_status=="Divorced" | mydata$marital_status=="Widowed" ~ "Separated"
)

mydata$overall_collapse <- rowSums(select(mydata, rash_collapse, hair_collapse, irritable_collapse, nausea_collapse))

mydata$bloodtest_collapse <- ifelse(mydata$blood_test_after_april_2014=="Yes", 1, 0)

mydata$phq <- rowSums(select(mydata, nervousscore, depressedscore, nopleasurescore, worryscore), na.rm=TRUE)
mydata$phq[is.na(mydata$nervousscore)] <- NA_real_

mydata$ptsd_bin <- as.numeric(mydata$ptsdscore >= 3)
mydata$phq_bin <- as.numeric(mydata$phq >= 6)

mydata$ptsd_bin <- factor(mydata$ptsd_bin, level=c(0,1), labels=c("No", "Yes"))
mydata$phq_bin <- factor(mydata$phq_bin, level=c(0,1), labels=c("No", "Yes"))

mydata$include <- ifelse(!is.na(mydata$age) & !is.na(mydata$reth) & !is.na(mydata$marital_status) & !is.na(mydata$educ_level) & !is.na(mydata$public_benefits) & !is.na(mydata$gender) & !is.na(mydata$confidants2) & !is.na(mydata$years_cityflint) & !is.na(mydata$current_zip) & !is.na(mydata$blood_test_after_april_2014) & !is.na(mydata$blood_test_before_april_2014) & !is.na(mydata$elevated_lead_before_april_2014) & !is.na(mydata$elevated_lead_after_april_2014) & !is.na(mydata$rashes_before_april_2014) & !is.na(mydata$rashes_after_april_2014) & !is.na(mydata$hairloss_before_april_2014) & !is.na(mydata$hairloss_after_april_2014) & !is.na(mydata$nausea_before_april_2014) & !is.na(mydata$nausea_after_april_2014) & !is.na(mydata$irritable_before_april_2014) & !is.na(mydata$irritable_after_april_2014) & !is.na(mydata$ptsdscore) & !is.na(mydata$phq) & mydata$gender != "Transgender Woman" & !is.na(mydata$num_time_notice),  1, 0)

mydata$include <- factor(mydata$include, levels=c(0, 1), labels=c("Excluded", "Included"))

mydata$employ_collapse <- case_when(
  mydata$employ_status=="Jobs/sporadic" | mydata$employ_status=="Other" ~ "Other",
  mydata$employ_status=="Retired" ~ "Retired",
  mydata$employ_status=="Employed" ~ "Employed",
  mydata$employ_status=="Unemployed" ~ "Unemployed",
  is.na(mydata$employ_status) ~ NA_character_)

var_label(mydata$collapseeduc) <- "Education level"
var_label(mydata$public_benefits) <- "Receives public benefits"
var_label(mydata$collapsemarital) <- "Marital status"
var_label(mydata$num_time_notice) <- "How long after water switch to notice water issue? (months)"
var_label(mydata$blood_test_after_april_2014) <- "Received blood lead test"
var_label(mydata$elevated_lead_after_april_2014) <- "Had elevated lead"
var_label(mydata$rashes_after_april_2014) <- "Had rashes"
var_label(mydata$hairloss_after_april_2014) <- "Had hair loss"
var_label(mydata$nausea_after_april_2014) <- "Had nausea"
var_label(mydata$irritable_after_april_2014) <- "Was irritable"
var_label(mydata$ptsd_bin) <- "PTSD"
var_label(mydata$phq_bin) <- "Depression/Anxiety"
var_label(mydata$employ_collapse) <- "Employment status"

mydata_complete <- filter(mydata, !is.na(age), !is.na(reth), !is.na(marital_status), !is.na(educ_level), !is.na(public_benefits), !is.na(gender), !is.na(confidants2), !is.na(years_cityflint), !is.na(current_zip), !is.na(blood_test_after_april_2014), !is.na(blood_test_before_april_2014), !is.na(elevated_lead_before_april_2014), !is.na(elevated_lead_after_april_2014), !is.na(rashes_before_april_2014), !is.na(rashes_after_april_2014), !is.na(hairloss_before_april_2014), !is.na(hairloss_after_april_2014), !is.na(nausea_before_april_2014), !is.na(nausea_after_april_2014), !is.na(irritable_before_april_2014), !is.na(irritable_after_april_2014), !is.na(ptsdscore), !is.na(phq), gender != "Transgender Woman", !is.na(num_time_notice), !is.na(employ_collapse))
```

We assemble some descriptives on the data:
```{r, echo=TRUE, message=FALSE, warning=FALSE}
table_predictors <- CreateTableOne(vars=c("age", "gender", "collapseeduc", "employ_collapse", "public_benefits", "collapsemarital", "number_children", "years_cityflint", "confidants2"), strata = "collapserace", data=mydata, includeNA = TRUE, argsNormal = list(NULL), argsNonNormal = list(var.equal = TRUE))

table_outcomes <- CreateTableOne(vars=c("num_time_notice", "blood_test_after_april_2014", "elevated_lead_after_april_2014", "rashes_after_april_2014", "hairloss_after_april_2014", "nausea_after_april_2014", "irritable_after_april_2014", "overall_collapse", "ptsd_bin", "phq_bin"), strata = "collapserace", data=mydata, argsNormal = list(NULL), argsNonNormal = list(var.equal = TRUE), includeNA = TRUE)

kable(print(table_predictors, contDigits=1, varLabel = T, printToggle=FALSE), booktabs=T) %>% add_indent(c(4:6, 8:12, 14:18, 20:22, 24:26)) %>% kable_styling(position="center")

kable(print(table_outcomes, contDigits=1, varLabel = T, printToggle=FALSE), booktabs=T) %>% add_indent(c(4:7, 9:12, 14:17, 19:22, 24:27, 29:32, 36:38)) %>% kable_styling(position="center")

table_predictors_notstrat <- CreateTableOne(vars=c("age", "gender", "collapseeduc", "employ_collapse", "public_benefits", "collapsemarital", "number_children", "years_cityflint", "confidants2"), data=mydata, argsNormal = list(NULL), argsNonNormal = list(var.equal = TRUE), includeNA=TRUE)

table_outcomes_notstrat <- CreateTableOne(vars=c("num_time_notice", "blood_test_after_april_2014", "elevated_lead_after_april_2014", "rashes_after_april_2014", "hairloss_after_april_2014", "nausea_after_april_2014", "irritable_after_april_2014", "overall_collapse", "ptsd_bin", "phq_bin"), data=mydata, argsNormal = list(NULL), argsNonNormal = list(var.equal = TRUE), includeNA=TRUE)

kable(print(table_predictors_notstrat, contDigits=1, varLabel = T, printToggle=FALSE), booktabs=T) %>% add_indent(c(4:6, 8:12, 14:18, 20:22, 24:26)) %>% kable_styling(position="center")

kable(print(table_outcomes_notstrat, contDigits=1, varLabel = T, printToggle=FALSE), booktabs=T) %>% add_indent(c(4:7, 9:12, 14:17, 19:22, 24:27, 29:32, 36:38)) %>% kable_styling(position="center")
```

We make maps and descriptive statistics:
```{r, echo=FALSE, message=FALSE, warning=FALSE}
zcta1 <- zctas(cb = TRUE, starts_with = c("4850", "48532"), class = "sf") %>% filter(ZCTA5CE10 != "48509")
flintdat <- data.frame("zip" = zcta1$ZCTA5CE10, "geometry" = zcta1$geometry)

fils <- unzip("~/Desktop/Research/Jerel Flint/Analyses/tl_2019_26_place.zip")
flint_boundary <- st_read(fils[4]) %>% filter(NAME=="Flint")

totalpop <- data.frame(table(mydata$current_zip))
colnames(totalpop) <- c("current_zip", "pop")

flintdat2 <- full_join(flintdat, totalpop, by=c("zip" = "current_zip")) 
sites <- read_csv("survey_locations.csv")

popmap <- ggplot(data=flintdat2) + geom_sf(data=flintdat2, aes(geometry = geometry, fill = factor(pop)), size=0, alpha = 0.6) + geom_sf(data=flint_boundary, aes(geometry=geometry), fill=NA, color = "black") + geom_sf_label(aes(geometry = geometry, label = paste0(zip, " (", pop, ")"))) + xlab(" ") + ylab(" ") + theme(rect = element_blank(), axis.ticks = element_blank(), axis.line=element_blank(), axis.text.x=element_blank(), axis.text.y=element_blank(), legend.position = "none") + scale_fill_grey() + geom_point(data = sites, aes(x = Longitude, y = Latitude), color = "red", shape = 8)

popmap

ggsave(filename = "~/Desktop/Research/Jerel Flint/Publications/Paper 1 - Adult Health/mymap.pdf", plot = popmap, width = 8, height = 8)

#table_predictors <- CreateTableOne(vars=c("age", "gender", "collapseeduc", "public_benefits", "collapsemarital", "number_children", "years_cityflint", "confidants2"), strata = "collapserace", data=mydata, includeNA = TRUE, argsNormal = list(NULL), argsNonNormal = list(var.equal = TRUE))

#table_outcomes <- CreateTableOne(vars=c("num_time_notice", "blood_test_after_april_2014", "elevated_lead_after_april_2014", "rashes_after_april_2014", "hairloss_after_april_2014", "nausea_after_april_2014", "irritable_after_april_2014", "overall_collapse", "ptsd_bin", "phq_bin"), strata = "collapserace", data=mydata, argsNormal = list(NULL), argsNonNormal = list(var.equal = TRUE), includeNA = TRUE)

#kable(print(table_predictors, contDigits=1, varLabel = T, printToggle=FALSE), booktabs=T) %>% add_indent(c(4:6, 8:12, 14:16, 18:20)) %>% kable_styling(position="center")

#kable(print(table_outcomes, contDigits=1, varLabel = T, printToggle=FALSE), booktabs=T) %>% add_indent(c(4:7, 9:12, 14:17, 19:22, 24:27, 29:32, 36:38)) %>% kable_styling(position="center")

#table_predictors_notstrat <- CreateTableOne(vars=c("age", "gender", "collapseeduc", "public_benefits", "collapsemarital", "number_children", "years_cityflint", "confidants2"), data=mydata, argsNormal = list(NULL), argsNonNormal = list(var.equal = TRUE), includeNA=TRUE)

#table_outcomes_notstrat <- CreateTableOne(vars=c("num_time_notice", "blood_test_after_april_2014", "elevated_lead_after_april_2014", "rashes_after_april_2014", "hairloss_after_april_2014", "nausea_after_april_2014", "irritable_after_april_2014", "overall_collapse", "ptsd_bin", "phq_bin"), data=mydata, argsNormal = list(NULL), argsNonNormal = list(var.equal = TRUE), includeNA=TRUE)

#kable(print(table_predictors_notstrat, contDigits=1, varLabel = T, printToggle=FALSE), booktabs=T) %>% add_indent(c(4:6, 8:12, 14:16, 18:20)) %>% kable_styling(position="center")

#kable(print(table_outcomes_notstrat, contDigits=1, varLabel = T, printToggle=FALSE), booktabs=T) %>% add_indent(c(4:7, 9:12, 14:17, 19:22, 24:27, 29:32, 36:38)) %>% kable_styling(position="center")
```

Now, we conduct Moran's I tests grouped by zipcode to reassure ourselves that we can just adjust for zipcode and not do a full-on spatial analysis:
```{r, echo=FALSE, message=FALSE, warning=FALSE}
zip_symp <- mydata_complete %>% dplyr::group_by(current_zip) %>% dplyr::summarize(elev_blood = sum(elevated_collapse, na.rm=TRUE), rash = sum(rash_collapse, na.rm=TRUE), nausea = sum(nausea_collapse, na.rm=TRUE), hairloss = sum(hair_collapse, na.rm=TRUE), irritable = sum(irritable_collapse, na.rm=TRUE), overall = sum(overall_collapse, na.rm=TRUE), ptsd = sum(ptsd_bin=="Yes", na.rm=TRUE), phq = sum(phq_bin=="Yes", na.rm=TRUE), get_test = sum(bloodtest_collapse, na.rm=TRUE), notice_issue = mean(num_time_notice, na.rm=TRUE), black = sum(collapserace=="Black", na.rm=TRUE), other = sum(collapserace=="Other", na.rm=TRUE),  l9 = sum(collapseeduc=="Less than high school", na.rm=TRUE), hs = sum(collapseeduc=="High school", na.rm=TRUE), assc = sum(collapseeduc=="Associate's", na.rm=TRUE), somecoll = sum(collapseeduc=="Some college", na.rm=TRUE), female = sum(gender!="Male", na.rm=TRUE), age = mean(age, na.rm=TRUE), pubben = sum(public_benefits=="Yes", na.rm=TRUE), pubbendk = sum(public_benefits=="Don't know/Unsure", na.rm=TRUE), network = mean(confidants2, na.rm=TRUE))

flintdat3 <- full_join(flintdat2, zip_symp, by=c("zip" = "current_zip"))

flintdat3$l9 <- flintdat3$l9/flintdat3$pop
flintdat3$hs <- flintdat3$hs/flintdat3$pop
flintdat3$assc <- flintdat3$assc/flintdat3$pop
flintdat3$somecoll <- flintdat3$somecoll/flintdat3$pop
flintdat3$black <- flintdat3$black/flintdat3$pop
flintdat3$other <- flintdat3$other/flintdat3$pop
flintdat3$female <- flintdat3$female/flintdat3$pop
flintdat3$pubben <- flintdat3$pubben/flintdat3$pop
flintdat3$pubbendk <- flintdat3$pubbendk/flintdat3$pop

flint.nb <- poly2nb(flintdat3$geometry)
flint.list.w <- nb2listw(flint.nb)

elevblood <- moran.test(flintdat3$elev_blood/flintdat3$pop, listw=flint.list.w, randomisation=FALSE)
elevblood
elevblood_adj <- lm.morantest(lm(flintdat3$elev_blood/flintdat3$pop~ flintdat3$black + flintdat3$other + flintdat3$female + flintdat3$age),listw=flint.list.w)
elevblood_adj

rash <- moran.test(flintdat3$rash/flintdat3$pop, listw=flint.list.w, randomisation=FALSE)
rash
rash_adj <- lm.morantest(lm(flintdat3$rash/flintdat3$pop ~ flintdat3$black + flintdat3$other + flintdat3$female + flintdat3$age),listw=flint.list.w)
rash_adj

nausea <- moran.test(flintdat3$nausea/flintdat3$pop, listw=flint.list.w, randomisation=FALSE)
nausea
nausea_adj <- lm.morantest(lm(flintdat3$nausea/flintdat3$pop ~ flintdat3$black + flintdat3$other + flintdat3$female + flintdat3$age),listw=flint.list.w)
nausea_adj

hairloss <- moran.test(flintdat3$hairloss/flintdat3$pop, listw=flint.list.w, randomisation=FALSE)
hairloss
hairloss_adj <- lm.morantest(lm(flintdat3$hairloss/flintdat3$pop ~ flintdat3$black + flintdat3$other + flintdat3$female + flintdat3$age),listw=flint.list.w)
hairloss_adj

irritable <- moran.test(flintdat3$irritable/flintdat3$pop, listw=flint.list.w, randomisation=FALSE)
irritable
irritable_adj <- lm.morantest(lm(flintdat3$irritable/flintdat3$pop ~ flintdat3$black + flintdat3$other + flintdat3$female + flintdat3$age),listw=flint.list.w)
irritable_adj

overall <- moran.test(flintdat3$overall/flintdat3$pop, listw=flint.list.w, randomisation=FALSE)
overall
overall_adj <- lm.morantest(lm(flintdat3$overall/flintdat3$pop ~ flintdat3$black + flintdat3$other + flintdat3$female + flintdat3$age),listw=flint.list.w)
overall_adj

ptsd <- moran.test(flintdat3$ptsd/flintdat3$pop, listw=flint.list.w, randomisation=FALSE)
ptsd
ptsd_adj <- lm.morantest(lm(flintdat3$ptsd/flintdat3$pop ~ flintdat3$black + flintdat3$other + flintdat3$female + flintdat3$age),listw=flint.list.w)
ptsd_adj

phq <- moran.test(flintdat3$phq/flintdat3$pop, listw=flint.list.w, randomisation=FALSE)
phq
phq_adj <- lm.morantest(lm(flintdat3$phq/flintdat3$pop ~ flintdat3$black + flintdat3$other + flintdat3$female + flintdat3$age),listw=flint.list.w)
phq_adj

get_tested <- moran.test(flintdat3$get_test/flintdat3$pop, listw=flint.list.w, randomisation=FALSE)
get_tested
get_tested_adj <- lm.morantest(lm(flintdat3$get_test/flintdat3$pop ~ flintdat3$black + flintdat3$other + flintdat3$female + flintdat3$age),listw=flint.list.w)
get_tested_adj

notice <- moran.test(flintdat3$notice_issue, listw=flint.list.w, randomisation=FALSE)
notice
notice_adj <- lm.morantest(lm(flintdat3$notice_issue ~ flintdat3$black + flintdat3$other + flintdat3$female + flintdat3$age),listw=flint.list.w)
notice_adj
```

There is some evidence of spatial correlation for nausea and hairloss, but once we adjust for baseline covariates it mostly goes away. Given the constraints of this analysis, we're going to just adjust for zipcode in our models and call it a day.

We fit our models according to the causal diagram given here:
```{r, echo=FALSE, message=FALSE, warning=FALSE}
knitr::include_graphics("~/Desktop/Research/Jerel Flint/Analyses/mediation.png")
```

Per reviewer request, we use multiple imputation with chained equations to impute missingness in the dataset:
```{r, echo=FALSE, message=FALSE, warning=FALSE}
mydata$collapsezip <- ifelse(mydata$current_zip==48502, 2, mydata$current_zip)
mydata$collapsezip <- factor(mydata$collapsezip, levels=c(2:7), labels=c("48502/48503", "48504", "48505", "48506", "48507", "48532"))
mydata$collapsepubben <- ifelse(mydata$public_benefits=="Yes", "Yes", "No")
mydata$collapsegender <- ifelse(mydata$gender=="Transgender Woman", "Female", mydata$gender)
mydata$rash_collapse_before <- ifelse(mydata$rashes_before_april_2014=="Yes", 1, 0)
mydata$hair_collapse_before <- ifelse(mydata$hairloss_before_april_2014=="Yes", 1, 0)
mydata$nausea_collapse_before <- ifelse(mydata$nausea_before_april_2014=="Yes", 1, 0)
mydata$irritable_collapse_before <- ifelse(mydata$irritable_before_april_2014=="Yes", 1, 0)
mydata$bloodtest_collapse_before <- ifelse(mydata$blood_test_before_april_2014=="Yes", 1, 0)

mydata_complete <- select(mydata, age, collapsegender, collapserace, collapsepubben, collapseeduc, employ_collapse, number_children, years_cityflint, confidants2, collapsezip, ptsd_bin, phq_bin, rash_collapse, hair_collapse, nausea_collapse, irritable_collapse, overall_collapse, elevated_collapse, bloodtest_collapse, num_time_notice, rash_collapse_before, hair_collapse_before, nausea_collapse_before, irritable_collapse_before, bloodtest_collapse_before)

mydata_complete$collapsegender <- factor(mydata_complete$collapsegender)
mydata_complete$collapserace <- factor(mydata_complete$collapserace)
mydata_complete$collapsepubben <- factor(mydata_complete$collapsepubben)
mydata_complete$collapseeduc <- factor(mydata_complete$collapseeduc)
mydata_complete$employ_collapse <- factor(mydata_complete$employ_collapse)
mydata_complete$collapsezip <- factor(mydata_complete$collapsezip)
mydata_complete$ptsd_bin <- factor(mydata_complete$ptsd_bin)
mydata_complete$phq_bin <- factor(mydata_complete$phq_bin)
mydata_complete$rash_collapse <- factor(mydata_complete$rash_collapse)
mydata_complete$hair_collapse <- factor(mydata_complete$hair_collapse)
mydata_complete$nausea_collapse <- factor(mydata_complete$nausea_collapse)
mydata_complete$irritable_collapse <- factor(mydata_complete$irritable_collapse)
mydata_complete$elevated_collapse <- factor(mydata_complete$elevated_collapse)
mydata_complete$bloodtest_collapse <- factor(mydata_complete$bloodtest_collapse)
mydata_complete$rash_collapse_before <- factor(mydata_complete$rash_collapse_before)
mydata_complete$hair_collapse_before <- factor(mydata_complete$hair_collapse_before)
mydata_complete$nausea_collapse_before <- factor(mydata_complete$nausea_collapse_before)
mydata_complete$irritable_collapse_before <- factor(mydata_complete$irritable_collapse_before)
mydata_complete$bloodtest_collapse_before <- factor(mydata_complete$bloodtest_collapse_before)

imp <- mice(mydata_complete, maxit = 0)

predM <- imp$predictorMatrix
meth <- imp$method

set.seed(123)
imp2 <- mice(mydata_complete, m = 25, print = FALSE)

imp_long <- mice::complete(imp2, action="long", include = TRUE)
imp_long$zip2 <- case_when(
  imp_long$collapsezip=="48502/48503" ~ "48502/48503",
  imp_long$collapsezip=="48504" ~ "48504",
  imp_long$collapsezip=="48505" ~ "48505",
  imp_long$collapsezip=="48506" ~ "48506",
  imp_long$collapsezip=="48507" | imp_long$collapsezip=="48532" ~ "48507/48532"
)

imp_long$collapseeduc2 <- case_when(
  imp_long$collapseeduc=="Less than high school" | imp_long$collapseeduc=="High school" ~ "High school or less",
  imp_long$collapseeduc=="Some college" ~ "Some college",
  imp_long$collapseeduc=="Associate's" | imp_long$collapseeduc=="College degree or more" ~ "Associate's or more"
)

imp_long$zip2 <- with(imp_long, as.factor(imp_long$zip2))
imp_long$collapseeduc2 <- with(imp_long, as.factor(imp_long$collapseeduc2))
imp_long_mids<-as.mids(imp_long)
```

As we know, when trying to understand the effect of a mediated effect on some outcome, we DO NOT adjust for the mediator. However, when trying to understand the effect of the mediator on the outcome, we DO adjust for the mediated effect. Therefore, we have to fit several models for our different outcomes, adjusting for different things depending on what the predictor of interest is. All told, we consider 10 different outcomes (PTSD, PHQ, rashes, hair loss, nausea, irritability, overall number of symptoms, elevated blood lead, getting tested, time to noticing issues) and consider the classes of predictors: demographics/baseline, time to noticing issues/getting tested, elevated blood lead, and emergence of symptoms. This results in us fitting 17 different models, given here:
```{r, echo=FALSE, message=FALSE, warning=FALSE}
effect_estims <- data.frame("Outcome" = NA, "Predictor" = NA, "Estimate" = NA, "SE" = NA, "Pval" = NA)

ptsd_dem <- with(imp_long_mids, glm(ptsd_bin ~ age + collapsegender + collapserace + collapsepubben + collapseeduc2 + employ_collapse + number_children + years_cityflint + confidants2 + zip2, family="binomial"))

summary(pool(ptsd_dem))

subdat <- data.frame("Outcome" = rep("PTSD", 17), "Predictor" = c("Age", "Male", "Other race", "White", "Receives benefits", "High school or less", "Some college", "Other employment", "Retired", "Unemployed", "Number of children", "Years in Flint", "Network size", "Zip 48504", "Zip 48505", "Zip 48506", "Zip 48507/48532"), "Estimate" = summary(pool(ptsd_dem))[-1,2], "SE" = summary(pool(ptsd_dem))[-1,3], "Pval" = summary(pool(ptsd_dem))[-1,6])

effect_estims <- rbind(effect_estims, subdat)

phq_dem <- with(imp_long_mids, glm(phq_bin ~ age + collapsegender + collapserace + collapsepubben + collapseeduc2 + employ_collapse + number_children + years_cityflint + confidants2 + zip2, family="binomial"))

summary(pool(phq_dem))

subdat <- data.frame("Outcome" = rep("PHQ", 17), "Predictor" = c("Age", "Male", "Other race", "White", "Receives benefits", "High school or less", "Some college", "Other employment", "Retired", "Unemployed",  "Number of children", "Years in Flint", "Network size", "Zip 48504", "Zip 48505", "Zip 48506", "Zip 48507/48532"), "Estimate" = summary(pool(phq_dem))[-1,2], "SE" = summary(pool(phq_dem))[-1,3], "Pval" = summary(pool(phq_dem))[-1,6])

effect_estims <- rbind(effect_estims, subdat)

rash_dem <- with(imp_long_mids, glm(rash_collapse ~ age + collapsegender + collapserace + collapsepubben + collapseeduc2 + employ_collapse + number_children + years_cityflint + confidants2 + zip2, family="binomial"))

summary(pool(rash_dem))

subdat <- data.frame("Outcome" = rep("Rashes", 17), "Predictor" = c("Age", "Male", "Other race", "White", "Receives benefits", "High school or less", "Some college", "Other employment", "Retired", "Unemployed", "Number of children", "Years in Flint", "Network size", "Zip 48504", "Zip 48505", "Zip 48506", "Zip 48507/48532"), "Estimate" = summary(pool(rash_dem))[-1,2], "SE" = summary(pool(rash_dem))[-1,3], "Pval" = summary(pool(rash_dem))[-1,6])

effect_estims <- rbind(effect_estims, subdat)

hairloss_dem <- with(imp_long_mids, glm(hair_collapse ~ age + collapsegender + collapserace + collapsepubben + collapseeduc2 + employ_collapse + number_children + years_cityflint + confidants2 + zip2, family="binomial"))

summary(pool(hairloss_dem))

subdat <- data.frame("Outcome" = rep("Hair loss", 17), "Predictor" = c("Age", "Male", "Other race", "White", "Receives benefits", "High school or less", "Some college", "Other employment", "Retired", "Unemployed", "Number of children", "Years in Flint", "Network size", "Zip 48504", "Zip 48505", "Zip 48506", "Zip 48507/48532"), "Estimate" = summary(pool(hairloss_dem))[-1,2], "SE" = summary(pool(hairloss_dem))[-1,3], "Pval" = summary(pool(hairloss_dem))[-1,6])

effect_estims <- rbind(effect_estims, subdat)

nausea_dem <- with(imp_long_mids, glm(nausea_collapse ~ age + collapsegender + collapserace + collapsepubben + collapseeduc2 + employ_collapse + number_children + years_cityflint + confidants2 + zip2, family="binomial"))

summary(pool(nausea_dem))

subdat <- data.frame("Outcome" = rep("Nausea", 17), "Predictor" = c("Age", "Male", "Other race", "White", "Receives benefits", "High school or less", "Some college", "Other employment", "Retired", "Unemployed", "Number of children", "Years in Flint", "Network size", "Zip 48504", "Zip 48505", "Zip 48506", "Zip 48507/48532"), "Estimate" = summary(pool(nausea_dem))[-1,2], "SE" = summary(pool(nausea_dem))[-1,3], "Pval" = summary(pool(nausea_dem))[-1,6])

effect_estims <- rbind(effect_estims, subdat)

irritable_dem <- with(imp_long_mids, glm(irritable_collapse ~ age + collapsegender + collapserace + collapsepubben + collapseeduc2 + employ_collapse + number_children + years_cityflint + confidants2 + zip2, family="binomial"))

summary(pool(irritable_dem))

subdat <- data.frame("Outcome" = rep("Agitation", 17), "Predictor" = c("Age", "Male", "Other race", "White", "Receives benefits", "High school or less", "Some college", "Other employment", "Retired", "Unemployed", "Number of children", "Years in Flint", "Network size", "Zip 48504", "Zip 48505", "Zip 48506", "Zip 48507/48532"), "Estimate" = summary(pool(irritable_dem))[-1,2], "SE" = summary(pool(irritable_dem))[-1,3], "Pval" = summary(pool(irritable_dem))[-1,6])

effect_estims <- rbind(effect_estims, subdat)

overall_dem <- with(imp_long_mids, glm(overall_collapse ~ age + collapsegender + collapserace + collapsepubben + collapseeduc2 + employ_collapse + number_children + years_cityflint + confidants2 + zip2, family="poisson"))

summary(pool(overall_dem))

subdat <- data.frame("Outcome" = rep("Overall", 17), "Predictor" = c("Age", "Male", "Other race", "White", "Receives benefits", "High school or less", "Some college", "Other employment", "Retired", "Unemployed", "Number of children", "Years in Flint", "Network size", "Zip 48504", "Zip 48505", "Zip 48506", "Zip 48507/48532"), "Estimate" = summary(pool(overall_dem))[-1,2], "SE" = summary(pool(overall_dem))[-1,3], "Pval" = summary(pool(overall_dem))[-1,6])

effect_estims <- rbind(effect_estims, subdat)

#bloodlead_dem <- with(imp_long_mids, logistf(elevated_collapse ~ age + collapsegender + collapserace + collapsepubben + number_children + years_cityflint + confidants2 + zip2))

#bloodlead_dem <- lapply(1:5, function(X) logistf(data = select(list_imps[[X]], elevated_collapse, age, collapsegender, collapserace, collapsepubben, number_children, years_cityflint, confidants2, zip2), elevated_collapse ~ age + collapsegender + collapserace + collapsepubben + number_children + years_cityflint + confidants2 + zip2, pl = TRUE, dataout = TRUE))

bloodlead_dem <- with(imp_long_mids, glm(elevated_collapse ~ age + collapsegender + collapserace + collapsepubben + collapseeduc2 + employ_collapse + number_children + years_cityflint + confidants2 + zip2, family = "binomial"))

#test1 <- summary(bloodlead_dem)

#test <- CLIP.confint(obj = bloodlead_dem)
summary(pool(bloodlead_dem))

subdat <- data.frame("Outcome" = rep("Elevated BLLs", 17), "Predictor" = c("Age", "Male", "Other race", "White", "Receives benefits", "High school or less", "Some college", "Other employment", "Retired", "Unemployed", "Number of children", "Years in Flint", "Network size", "Zip 48504", "Zip 48505", "Zip 48506", "Zip 48507/48532"), "Estimate" = summary(pool(bloodlead_dem))[-1,2], "SE" = summary(pool(bloodlead_dem))[-1,3], "Pval" = summary(pool(bloodlead_dem))[-1,6])

effect_estims <- rbind(effect_estims, subdat)

gettest_dem <- with(imp_long_mids, glm(bloodtest_collapse ~ age + collapsegender + collapserace + collapsepubben + collapseeduc2 + employ_collapse + number_children + years_cityflint + confidants2 + zip2, family="binomial"))

summary(pool(gettest_dem))

subdat <- data.frame("Outcome" = rep("Get tested", 17), "Predictor" = c("Age", "Male", "Other race", "White", "Receives benefits", "High school or less", "Some college", "Other employment", "Retired", "Unemployed", "Number of children", "Years in Flint", "Network size", "Zip 48504", "Zip 48505", "Zip 48506", "Zip 48507/48532"), "Estimate" = summary(pool(gettest_dem))[-1,2], "SE" = summary(pool(gettest_dem))[-1,3], "Pval" = summary(pool(gettest_dem))[-1,6])

effect_estims <- rbind(effect_estims, subdat)

time_dem <- with(imp_long_mids, lm(log(num_time_notice) ~ age + collapsegender + collapserace + collapsepubben + collapseeduc2 + employ_collapse + number_children + years_cityflint + confidants2 + zip2))

summary(pool(time_dem))

subdat <- data.frame("Outcome" = rep("Time to notice", 17), "Predictor" = c("Age", "Male", "Other race", "White", "Receives benefits", "High school or less", "Some college", "Other employment", "Retired", "Unemployed", "Number of children", "Years in Flint", "Network size", "Zip 48504", "Zip 48505", "Zip 48506", "Zip 48507/48532"), "Estimate" = summary(pool(time_dem))[-1,2], "SE" = summary(pool(time_dem))[-1,3], "Pval" = summary(pool(time_dem))[-1,6])

effect_estims <- rbind(effect_estims, subdat)

ptsd_symp <- with(imp_long_mids, glm(ptsd_bin ~ age + collapsegender + collapserace + collapsepubben + collapseeduc2 + employ_collapse + number_children + years_cityflint + confidants2 + zip2 + overall_collapse + elevated_collapse + bloodtest_collapse + num_time_notice, family="binomial"))

summary(pool(ptsd_symp))

subdat <- data.frame("Outcome" = rep("PTSD", 4), "Predictor" = c("Number of symptoms", "Elevated BLLs", "Get tested", "Time to notice"), "Estimate" = summary(pool(ptsd_symp))[19:22,2], "SE" = summary(pool(ptsd_symp))[19:22,3], "Pval" = summary(pool(ptsd_symp))[19:22,6])

effect_estims <- rbind(effect_estims, subdat)

phq_symp <- with(imp_long_mids, glm(phq_bin ~ age + collapsegender + collapserace + collapsepubben + collapseeduc2 + employ_collapse + number_children + years_cityflint + confidants2 + zip2 + overall_collapse + elevated_collapse + bloodtest_collapse + num_time_notice, family="binomial"))

summary(pool(phq_symp))

subdat <- data.frame("Outcome" = rep("PHQ", 4), "Predictor" = c("Number of symptoms", "Elevated BLLs", "Get tested", "Time to notice"), "Estimate" = summary(pool(phq_symp))[19:22,2], "SE" = summary(pool(phq_symp))[19:22,3], "Pval" = summary(pool(phq_symp))[19:22,6])

effect_estims <- rbind(effect_estims, subdat)

#rash_symp <- logistf(rash_collapse ~ age + gender + collapserace + collapsepubben + number_children + years_cityflint + confidants2 + zip2 + hair_collapse + nausea_collapse + irritable_collapse + elevated_collapse + bloodtest_collapse + num_time_notice + rash_collapse_before, data=mydata_complete)

#test2 <- summary(rash_symp)

rash_symp <- with(imp_long_mids, glm(rash_collapse ~ age + collapsegender + collapserace + collapsepubben + collapseeduc2 + employ_collapse + number_children + years_cityflint + confidants2 + zip2 + hair_collapse + nausea_collapse + irritable_collapse + elevated_collapse + bloodtest_collapse + num_time_notice + rash_collapse_before, family = "binomial"))

summary(pool(rash_symp))

subdat <- data.frame("Outcome" = rep("Rashes", 7), "Predictor" = c("Hair loss", "Nausea", "Agitation", "Elevated BLLs", "Get tested", "Time to notice", "Rashes before"), "Estimate" = summary(pool(rash_symp))[19:25,2], "SE" = summary(pool(rash_symp))[19:25,3], "Pval" = summary(pool(rash_symp))[19:25,6])

effect_estims <- rbind(effect_estims, subdat)

#hairloss_symp <- logistf(hair_collapse ~ age + gender + collapserace + collapsepubben + number_children + years_cityflint + confidants2 + zip2 + rash_collapse + nausea_collapse + irritable_collapse + elevated_collapse + bloodtest_collapse + num_time_notice + hair_collapse_before, data=mydata_complete)

#test3 <- summary(hairloss_symp)

hairloss_symp <- with(imp_long_mids, glm(hair_collapse ~ age + collapsegender + collapserace + collapsepubben + collapseeduc2 + employ_collapse + number_children + years_cityflint + confidants2 + zip2 + rash_collapse + nausea_collapse + irritable_collapse + elevated_collapse + bloodtest_collapse + num_time_notice + hair_collapse_before, family = "binomial"))

summary(pool(hairloss_symp))

subdat <- data.frame("Outcome" = rep("Hair loss", 7), "Predictor" = c("Rashes", "Nausea", "Agitation", "Elevated BLLs", "Get tested", "Time to notice", "Hair loss before"), "Estimate" = summary(pool(hairloss_symp))[19:25,2], "SE" = summary(pool(hairloss_symp))[19:25,3], "Pval" = summary(pool(hairloss_symp))[19:25,6])

effect_estims <- rbind(effect_estims, subdat)

#nausea_symp <- logistf(nausea_collapse ~ age + gender + collapserace + collapsepubben + number_children + years_cityflint + confidants2 + zip2 + rash_collapse + hair_collapse + irritable_collapse + elevated_collapse + bloodtest_collapse + num_time_notice + nausea_collapse_before, data=mydata_complete)

#test4 <- summary(nausea_symp)

nausea_symp <- with(imp_long_mids, glm(nausea_collapse ~ age + collapsegender + collapserace + collapsepubben + collapseeduc2 + employ_collapse + number_children + years_cityflint + confidants2 + zip2 + rash_collapse + hair_collapse + irritable_collapse + elevated_collapse + bloodtest_collapse + num_time_notice + nausea_collapse_before, family = "binomial"))

summary(pool(nausea_symp))

subdat <- data.frame("Outcome" = rep("Nausea", 7), "Predictor" = c("Rashes", "Hair loss", "Agitation", "Elevated BLLs", "Get tested", "Time to notice", "Nausea before"), "Estimate" = summary(pool(nausea_symp))[19:25,2], "SE" = summary(pool(nausea_symp))[19:25,3], "Pval" = summary(pool(nausea_symp))[19:25,6])

effect_estims <- rbind(effect_estims, subdat)

#irritable_symp <- logistf(irritable_collapse ~ age + gender + collapserace + collapsepubben + number_children + years_cityflint + confidants2 + zip2 + rash_collapse + hair_collapse + nausea_collapse + elevated_collapse + bloodtest_collapse + num_time_notice + irritable_collapse_before, data=mydata_complete)

#test5 <- summary(irritable_symp)

irritable_symp <- with(imp_long_mids, glm(irritable_collapse ~ age + collapsegender + collapserace + collapsepubben + collapseeduc2 + employ_collapse + number_children + years_cityflint + confidants2 + zip2 + rash_collapse + hair_collapse + nausea_collapse + elevated_collapse + bloodtest_collapse + num_time_notice + irritable_collapse_before, family = "binomial"))

summary(pool(irritable_symp))

subdat <- data.frame("Outcome" = rep("Agitation", 7), "Predictor" = c("Rashes", "Hair loss", "Nausea", "Elevated BLLs", "Get tested", "Time to notice", "Agitation before"), "Estimate" = summary(pool(irritable_symp))[19:25,2], "SE" = summary(pool(irritable_symp))[19:25,3], "Pval" = summary(pool(irritable_symp))[19:25,6])

effect_estims <- rbind(effect_estims, subdat)

overall_bloodlead <- with(imp_long_mids, glm(overall_collapse ~ age + collapsegender + collapserace + collapsepubben + collapseeduc2 + employ_collapse + number_children + years_cityflint + confidants2 + zip2 + elevated_collapse + bloodtest_collapse + num_time_notice + rash_collapse_before + hair_collapse_before + nausea_collapse_before + irritable_collapse_before, family = "poisson"))

summary(pool(overall_bloodlead))

subdat <- data.frame("Outcome" = rep("Overall symptoms", 7), "Predictor" = c("Elevated BLLs", "Get tested", "Time to notice", "Rashes before", "Hair loss before", "Nausea before", "Agitation before"), "Estimate" = summary(pool(overall_bloodlead))[19:25,2], "SE" = summary(pool(overall_bloodlead))[19:25,3], "Pval" = summary(pool(overall_bloodlead))[19:25,6])

effect_estims <- rbind(effect_estims, subdat)
```

Since we fit quite a few models, we use the FDR to adjust for our multiple testing.
```{r, echo=FALSE, message=FALSE, warning=FALSE}
effect_estims <- effect_estims[-1,]
effect_estims$Lower <- effect_estims$Estimate - 1.96*effect_estims$SE
effect_estims$Upper <- effect_estims$Estimate + 1.96*effect_estims$SE

pretty_effects <- data.frame("Outcome" = effect_estims$Outcome, "Predictor" = effect_estims$Predictor, "Estimate" = exp(effect_estims$Estimate), "Lower" = exp(effect_estims$Lower), "Upper" = exp(effect_estims$Upper), "BHpvalue" = round(p.adjust(effect_estims$Pval, method="BH"), digits = 2))

pretty_effects$Label <- paste0(round(pretty_effects$Estimate, digits = 1), " (", round(pretty_effects$Lower, digits = 1), ", ", round(pretty_effects$Upper, digits = 1), ")")

pretty_effects2 <- select(pretty_effects, Outcome, Predictor, Label, BHpvalue)
#ptsd_plot <- ggplot(myresults[myresults$Outcome=="PTSD",], aes(fct_rev(Predictor_clean), Coefficient)) + geom_point() + geom_errorbar(aes(ymin=Lower, ymax=Upper, width=0.1)) + coord_flip(ylim=c(-2.35, 5.5)) + labs(y = "Odds Ratio and 95% CI", x = "", caption=TeX('Protective $\\leftarrow \\rightarrow$ Harmful' )) + geom_hline(yintercept=0, linetype="dashed", color = "black") + theme(panel.background = element_rect(fill="white"), panel.grid.major = element_blank(), panel.grid.minor=element_blank(), legend.position="bottom", text=element_text(size=14), plot.caption = element_text(size=14, hjust = 0.15)) + geom_rect(fill="white", ymin=4.7, ymax=Inf, xmin=-Inf, xmax=Inf) + annotate(geom = 'text', label = myresults$Label[myresults$Outcome=="PTSD"], x = myresults$Predictor_clean[myresults$Outcome=="PTSD"], y = 3.5, hjust = 0, vjust = 0.5) + scale_y_continuous(breaks=log(c(0.1, 0.25, 0.5, 1, 2, 4, 10)), labels=c("0.1","0.25", "0.5", "1.0", "2.0", "4.0", "10.0")) + ggtitle("Outcome: Trauma Symptoms")

#ptsd_plot

#phq_plot <- ggplot(myresults[myresults$Outcome=="PHQ",], aes(fct_rev(Predictor_clean), Coefficient)) + geom_point() + geom_errorbar(aes(ymin=Lower, ymax=Upper, width=0.1)) + coord_flip(ylim=c(-2.35, 5.5)) + labs(y = "Odds Ratio and 95% CI", x = "", caption=TeX('Protective $\\leftarrow \\rightarrow$ Harmful' )) + geom_hline(yintercept=0, linetype="dashed", color = "black") + theme(panel.background = element_rect(fill="white"), panel.grid.major = element_blank(), panel.grid.minor=element_blank(), legend.position="bottom", text=element_text(size=14), plot.caption = element_text(size=14, hjust = 0.15), axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank()) + geom_rect(fill="white", ymin=4.7, ymax=Inf, xmin=-Inf, xmax=Inf) + annotate(geom = 'text', label = myresults$Label[myresults$Outcome=="PHQ"], x = myresults$Predictor_clean[myresults$Outcome=="PHQ"], y = 3.5, hjust = 0, vjust = 0.5) + scale_y_continuous(breaks=log(c(0.1, 0.25, 0.5, 1, 2, 4, 10)), labels=c("0.1","0.25", "0.5", "1.0", "2.0", "4.0", "10.0")) + ggtitle("Outcome: Depression/Anxiety")

#phq_plot

#mental_plots <- grid.arrange(ptsd_plot, phq_plot, nrow=1, widths = c(6, 3))

```