### 2021-10-05 paperfour_healthbehaviors.Rmd
Update to include analysis for health behavior paper, and to revise child health analysis using Bayesian models.

---
title: 'Manuscript 4: Health Behaviors after the Crisis'
author: "Elizabeth Chase"
date: "5/28/2021"
output: pdf_document
---

```{r setup, include=FALSE, echo=TRUE}
library(knitr)
opts_chunk$set(collapse=TRUE, fig.align='center', out.extra = '', fig.pos= "h", tidy=TRUE, tidy.opts=list(blank=TRUE, width.cutoff=40), warning=FALSE,message=FALSE)
options(scipen=999)

rm(list=ls())

library(tidyverse)
library(gridExtra)
library(RColorBrewer)
library(tableone)
library(latex2exp)
library(labelled)
library(kableExtra)
library(logistf)
library(mice)
library(survival)
library(survminer)
library(MASS)
library(ggalluvial)

mydata <- readRDS("~/Desktop/Research/Jerel Flint/Analyses/adult_data_clean.rds")

load("~/Desktop/Research/Jerel Flint/Analyses/child_long.RData")

child_sub <- group_by(child_final, unique_id) %>% summarize(bottledwater_child = sum(as.numeric(watersource=="Bottled water"), na.rm = TRUE), testpost2014_child = sum(as.numeric(blood_test_after_april_2014=="Yes"), na.rm = TRUE), bllpost2014_child = sum(as.numeric(elevated_lead_after_april_2014=="Yes"), na.rm = TRUE))

mydata <- left_join(mydata, child_sub, by = c("unique_id"))

mydata$bottledwater_child[mydata$number_children == 0] <- 0 
mydata$testpost2014_child[mydata$number_children == 0] <- 0 
mydata$bllpost2014_child[mydata$number_children == 0] <- 0

mydata$num_time_notice <- case_when(
  mydata$noticed_water_issue=="1 year *AFTER* it was switched" ~ 12,
  mydata$noticed_water_issue=="2 months *AFTER* it was switched" ~ 2,
  mydata$noticed_water_issue=="Unsure/Don't know" ~ NA_real_,
  mydata$noticed_water_issue=="2 years *AFTER* it was switched" ~ 24,
  mydata$noticed_water_issue=="Less than 1 month *AFTER* the water source was switched" ~ 0.5,
  mydata$noticed_water_issue=="1 month *AFTER* it was switched" ~ 1,
  mydata$noticed_water_issue=="3 months *AFTER* it was switched" ~ 3,
  mydata$noticed_water_issue=="1 year and 2 months *AFTER* it was switched" ~ 14,
  mydata$noticed_water_issue=="6 months *AFTER* it was switched" ~ 6,
  mydata$noticed_water_issue=="9 months *AFTER* it was switched" ~ 9,
  mydata$noticed_water_issue=="1 year and 6 months *AFTER* it was switched" ~ 18,
  mydata$noticed_water_issue=="7 months *AFTER* it was switched" ~ 7,
  mydata$noticed_water_issue=="4 years or more *AFTER* it was switched" ~ 4,
  mydata$noticed_water_issue=="10 months *AFTER* it was switched" ~ 10,
  mydata$noticed_water_issue=="5 months *AFTER* it was switched" ~ 5,
  mydata$noticed_water_issue=="11 months *AFTER* it was switched" ~ 11,
  mydata$noticed_water_issue=="4 months *AFTER* it was switched" ~ 4,
  mydata$noticed_water_issue=="1 year and 5 months *AFTER* it was switched" ~ 17,
  mydata$noticed_water_issue=="8 months *AFTER* it was switched" ~ 8,
  mydata$noticed_water_issue=="3 years *AFTER* it was switched" ~ 36,
  mydata$noticed_water_issue=="2 year and 1 month *AFTER* it was switched" ~ 25,
  mydata$noticed_water_issue=="2 year and 6 months *AFTER* it was switched" ~ 30,
  mydata$noticed_water_issue=="1 year and 3 months *AFTER* it was switched" ~ 15,
  mydata$noticed_water_issue=="1 year and 1 month *AFTER* it was switched" ~ 13
)

mydata$notice_date <- as.Date("2014-04-25") + (mydata$num_time_notice*30)

mydata$long_ago_test[mydata$watertestnum==0 & is.na(mydata$long_ago_test)] <- 0
mydata$water_test_event_ind <- ifelse(!is.na(mydata$long_ago_test), 1, ifelse(mydata$watertestnum==0 & is.na(mydata$long_ago_test), 0, NA_real_))

mydata$last_test_date <- mydata$survey_date - (mydata$long_ago_test*30)
mydata$water_test_surv <- round(as.numeric(mydata$last_test_date - as.Date("2014-04-25"))/30, digits = 0)
mydata$water_test_event_ind[mydata$water_test_surv < 0 & !is.na(mydata$water_test_surv)] <- 0
mydata$water_test_surv[mydata$water_test_surv < 0 & !is.na(mydata$water_test_surv)] <- round(as.numeric(mydata$survey_date[mydata$water_test_surv < 0 & !is.na(mydata$water_test_surv)] - as.Date("2014-04-25"))/30, digits = 0)

mydata$waterlinereplaced_longago_num <- case_when(
  mydata$waterline_replaced_when=="1 month ago" ~ 1,
  mydata$waterline_replaced_when=="1 year ago" ~ 12,
  mydata$waterline_replaced_when=="1 year and 10 months ago" ~ 22,
  mydata$waterline_replaced_when=="1 year and 2 months ago" ~ 14,
  mydata$waterline_replaced_when=="1 year and 3 months ago" ~ 15,
  mydata$waterline_replaced_when=="1 year and 4 months ago" ~ 16,
  mydata$waterline_replaced_when=="1 year and 6 months ago" ~ 18,
  mydata$waterline_replaced_when=="1 year and 7 months ago" ~ 19,
  mydata$waterline_replaced_when=="10 months ago" ~ 10,
  mydata$waterline_replaced_when=="2 months ago" ~ 2,
  mydata$waterline_replaced_when=="2 years ago" ~ 24,
  mydata$waterline_replaced_when=="2 years and 1 month ago" ~ 25,
  mydata$waterline_replaced_when=="2 years and 3 months ago" ~ 27,
  mydata$waterline_replaced_when=="2 years and 4 months ago" ~ 28,
  mydata$waterline_replaced_when=="2 years and 6 months ago" ~ 30,
  mydata$waterline_replaced_when=="2 years and 7 months ago" ~ 31,
  mydata$waterline_replaced_when=="3 months ago" ~ 3,
  mydata$waterline_replaced_when=="3 years and 6 months ago" ~ 42,
  mydata$waterline_replaced_when=="4 months ago" ~ 4,
  mydata$waterline_replaced_when=="4 years and 6 months ago" ~ 54,
  mydata$waterline_replaced_when=="5 months ago" ~ 5,
  mydata$waterline_replaced_when=="6 months ago" ~ 6,
  mydata$waterline_replaced_when=="6 OR MORE YEARS AGO" ~ 72,
  mydata$waterline_replaced_when=="7 months ago" ~ 7,
  mydata$waterline_replaced_when=="8 months ago" ~ 8,
  mydata$waterline_replaced_when=="Less than 1 month ago" ~ 0.5,
)

mydata$waterline_cured_ind <- ifelse(mydata$water_line_status_after_april2014=="A mailing or official said that my lead/water line did not need to be replaced" | mydata$water_line_status_after_april2014=="I don't live in a home" | mydata$water_line_status_after_april2014=="I know my lead/water line didn't need to be replaced but wasn't told this in a mailing or by an official", 1, ifelse(mydata$water_line_status_after_april2014=="I don't know what the status is" | mydata$water_line_status_after_april2014=="Other" | is.na(mydata$water_line_status_after_april2014), NA_real_, 0))

mydata$waterline_replaced_eventind <- ifelse(!is.na(mydata$waterlinereplaced_longago_num) | mydata$water_line_status_after_april2014=="My lead/water lines were replaced", 1, ifelse(mydata$waterline_cured_ind==1 | mydata$water_line_status_after_april2014=="I was offered a lead/water line replacement but I refused it" | mydata$water_line_status_after_april2014=="My lead/water lines haven't been replaced but I was told in a mailing/by an official they will be", 0, NA_real_))

mydata$waterlinereplaced_longago_num[mydata$waterline_replaced_eventind==0] <- 0

mydata$waterline_date <- mydata$survey_date - (mydata$waterlinereplaced_longago_num*30)

mydata$water_line_surv <- round(as.numeric(mydata$waterline_date - as.Date("2014-04-25"))/30, digits = 0)
mydata$waterline_replaced_eventind[mydata$water_line_surv < 0 & !is.na(mydata$water_line_surv)] <- 0
mydata$water_line_surv[mydata$water_line_surv < 0 & !is.na(mydata$water_line_surv)] <- round(as.numeric(mydata$survey_date[mydata$water_line_surv < 0 & !is.na(mydata$water_line_surv)] - as.Date("2014-04-25"))/30, digits = 0)
```

We start with our own data exploration: 
```{r, echo=FALSE, message=FALSE, warning=FALSE}
print("Age:")
quantile(mydata$age)
ggplot(data = mydata, aes(x = age)) + geom_histogram() + theme_bw() + xlab("Age")

print("Gender:")
table(mydata$gender, useNA = "ifany")

print("Race/Ethnicity:")
table(mydata$reth, useNA = "ifany")

print("Education:")
table(mydata$educ_level, useNA = "ifany")

print("Marital status:")
table(mydata$marital_status, useNA = "ifany")

print("Employment status:")
table(mydata$employ_status, useNA = "ifany")

print("Years lived in Flint:")
quantile(mydata$years_cityflint)
ggplot(data = mydata, aes(x = years_cityflint)) + geom_histogram() + theme_bw() + xlab("Years in Flint")

print("Current housing:")
table(mydata$current_housing, useNA = "ifany")

print("Current ZIP code:")
table(mydata$current_zip, useNA = "ifany")

print("Health insurance:")
table(mydata$have_health_insurance, useNA = "ifany")

print("Vehicle access:")
table(mydata$access_vehicle, useNA = "ifany")

print("Water source:")
table(mydata$water_source, useNA = "ifany")

print("Months to notice a problem:")
print(paste0("Number missing: ", length(which(is.na(mydata$num_time_notice)))))
quantile(mydata$num_time_notice, na.rm = TRUE)
ggplot(data = mydata, aes(x = num_time_notice)) + geom_histogram() + theme_bw() + xlab("Months to notice a problem")

print("How many times waterline tested:")
print(paste0("Number missing: ", length(which(is.na(mydata$watertestnum)))))
quantile(mydata$watertestnum, na.rm = TRUE)
ggplot(data = mydata, aes(x = watertestnum)) + geom_histogram() + theme_bw() + xlab("Number of times tested")

print("Time of last water line test:")
print(paste0("Number missing: ", length(which(is.na(mydata$water_test_event_ind)))))
ggplot(data = mydata, aes(x = last_test_date)) + geom_bar() + theme_bw() + xlab("Last test date")
km_watertest <- survfit(Surv(water_test_surv, water_test_event_ind) ~ 1, data = mydata, conf.type = "log-log")
ggsurvplot(km_watertest, data=mydata, conf.int=TRUE, risk.table=TRUE, xlab = "Time (months)", 
                                title = "Time to Last Water Test")

print("Last water line test result:")
table(mydata$collapse_last_test, useNA = "ifany")

print("Adult blood test before April 2014:")
table(mydata$blood_test_before_april_2014, useNA = "ifany")

print("Adult blood test after April 2014:")
table(mydata$blood_test_after_april_2014, useNA = "ifany")

print("Elevated BLLs before April 2014:")
table(mydata$elevated_lead_before_april_2014, useNA = "ifany")

print("Elevated BLLs after April 2014:")
table(mydata$elevated_lead_after_april_2014, useNA = "ifany")

print("Current water source:")
table(mydata$current_water_source, useNA = "ifany")

print("Bottled water for drinking:")
table(mydata$bottled_water_drink, useNA = "ifany")

print("Bottled water for cooking:")
table(mydata$bottled_water_cook, useNA = "ifany")

print("Bottled water for bathing:")
table(mydata$bottled_water_bathe, useNA = "ifany")

print("Bottled water for brushing teeth:")
table(mydata$bottled_water_teeth, useNA = "ifany")

print("Bottled water for cleaning:")
table(mydata$bottled_water_cleaning, useNA = "ifany")

print("Bottled water for other purpose:")
table(mydata$bottled_water_other, useNA = "ifany")

print("Number of bottles per day:")
quantile(mydata$num_bottles_per_day, na.rm = TRUE)
print(paste0("Number missing: ", length(which(is.na(mydata$num_bottles_per_day)))))
ggplot(data = mydata, aes(x = num_bottles_per_day)) + geom_histogram() + theme_bw() + xlab("Number of bottles used per day")

print("Bottle disposal habits:")
table(mydata$water_bottle_disposal, useNA = "ifany")

print("Water line status after April 2014:")
table(mydata$water_line_status_after_april2014, useNA = "ifany")

print("Water line replaced when:")
print(paste0("Number missing: ", length(which(is.na(mydata$waterline_replaced_eventind)))))
ggplot(data = mydata, aes(x = waterline_date)) + geom_bar() + theme_bw() + xlab("Date of waterline replacement")
km_waterline <- survfit(Surv(water_line_surv, waterline_replaced_eventind) ~ 1, data = mydata[mydata$waterline_cured_ind==0,], conf.type = "log-log")
ggsurvplot(km_waterline, data=mydata[mydata$waterline_cured_ind==0,], conf.int=TRUE, risk.table=TRUE, xlab = "Time (months)", title = "Time to Waterline Replacement")

print("Ever used water pod:")
table(mydata$used_water_pod, useNA = "ifany")

print("Currently use water pod:")
table(mydata$get_free_bottled_water, useNA = "ifany")

print("Don't get bottled water because didn't know about it:")
table(mydata$no_free_bottles_dont_know, useNA = "ifany")

print("Don't get bottled water because of supply limits:")
table(mydata$no_free_bottles_supply_limit, useNA = "ifany")

print("Don't get bottled water because they're too far:")
table(mydata$no_free_bottles_too_far, useNA = "ifany")

print("Don't get bottled water because I lack transportation:")
table(mydata$no_free_bottles_no_transport, useNA = "ifany")

print("Don't get bottled water because their hours are bad:")
table(mydata$no_free_bottles_bad_hours, useNA = "ifany")

print("Don't get bottled water because the lines are long:")
table(mydata$no_free_bottles_long_lines, useNA = "ifany")

print("Don't get bottled water because I use my tap water:")
table(mydata$no_free_bottles_use_tap, useNA = "ifany")

print("Don't get bottled water because of another reason:")
table(mydata$no_free_bottles_other, useNA = "ifany")

print("Bottled water site:")
table(mydata$bottle_site, useNA = "ifany")

print("Will drink tap water in three years:")
table(mydata$drink_tap_in_three_years, useNA = "ifany")

print("Social network size:")
quantile(mydata$confidants2)
ggplot(data = mydata, aes(x = confidants2)) + geom_histogram() + theme_bw() + xlab("Social network size")

print("Number of children in care:")
quantile(mydata$number_children)
ggplot(data = mydata, aes(x = number_children)) + geom_histogram() + theme_bw() + xlab("Number of children in care")

print("Percent of children in care drinking bottled water:")
table(mydata$bottledwater_child/mydata$number_children, useNA = "ifany")
ggplot(data = mydata, aes(x = bottledwater_child/number_children)) + geom_bar() + theme_bw() + xlab("Percent of children drinking bottled water")

print("Percent of children in care tested for BLLs post-2014:")
table(mydata$testpost2014_child/mydata$number_children, useNA = "ifany")
ggplot(data = mydata, aes(x = testpost2014_child/number_children)) + geom_bar() + theme_bw() + xlab("Percent of children tested for BLLs after 2014")

print("Percent of children in care with elevated BLLs post-2014:")
table(mydata$bllpost2014_child/mydata$number_children, useNA = "ifany")
ggplot(data = mydata, aes(x = bllpost2014_child/number_children)) + geom_bar() + theme_bw() + xlab("Percent of children with elevated BLLs after 2014")

print("Child age:")
quantile(child_final$age)
ggplot(data = child_final, aes(x = age)) + geom_histogram() + theme_bw() + xlab("Child age")

print("Child gender:")
table(child_final$gender, useNA = "ifany")

print("Child race:")
table(child_final$reth, useNA = "ifany")

print("Child water source:")
table(child_final$watersource, useNA = "ifany")

print("Child tested for BLLs pre-2014:")
table(child_final$blood_test_before_april_2014, useNA = "ifany")

print("Child tested for BLLs post-2014:")
table(child_final$blood_test_after_april_2014, useNA = "ifany")

print("Child had elevated BLLs pre-2014:")
table(child_final$elevated_lead_before_april_2014, useNA = "ifany")

print("Child had elevated BLLs post-2014:")
table(child_final$elevated_lead_after_april_2014, useNA = "ifany")
```

Now, we do the multiple imputation we'll need for modeling further down the line:
```{r, echo=FALSE, message=FALSE, warning=FALSE, cache = TRUE}
vars<-c("age",
        "gender",
        "reth",
        "educ_level",
        "public_benefits",
        "employ_status",
        "marital_status",
        "years_cityflint",
        "current_housing",
        "current_zip",
        "have_health_insurance",
        "access_vehicle",
        "water_source",
        "num_time_notice",
        "watertestnum",
        "water_test_surv", 
        "water_test_event_ind",
        "collapse_last_test",
        "blood_test_before_april_2014",
        "blood_test_after_april_2014",
        "elevated_lead_before_april_2014",
        "elevated_lead_after_april_2014",
        "current_water_source",
        "bottled_water_drink",
        "bottled_water_cook",
        "bottled_water_bathe",
        "bottled_water_teeth",
        "bottled_water_cleaning",
        "bottled_water_other",
        "num_bottles_per_day",
        "number_children",
        "water_bottle_disposal",
        "water_line_status_after_april2014",
        "water_line_surv", 
        "waterline_replaced_eventind",
        "waterline_cured_ind",
        "used_water_pod",
        "get_free_bottled_water",
        "no_free_bottles_dont_know",
        "no_free_bottles_supply_limit",
        "no_free_bottles_too_far",
        "no_free_bottles_no_transport",
        "no_free_bottles_bad_hours",
        "no_free_bottles_long_lines",
        "no_free_bottles_use_tap",
        "no_free_bottles_other",
        "bottle_site",
        "drink_tap_in_three_years",
        "confidants2",
        "bllpost2014_child",
        "testpost2014_child",
        "bottledwater_child"
        )

missing_plot <- VIM::aggr(mydata[,vars],
                          numbers = T,
                          sortVars = T,
                          cex.lab=0.8,
                          cex.axis=0.8,
                          # col = c("lightgreen", "darkred", "orange"),
                          # labels=str_sub(names(train[,vars]), 1, 8),
                          ylab=c("Missing Value Counts"
                                 , "Pattern"))
missing_plot

mydata_complete <- mydata[,vars]

#mydata_complete$gender <- factor(mydata_complete$gender)
mydata_complete$reth <- factor(mydata_complete$reth)
mydata_complete$educ_level <- factor(mydata_complete$educ_level)
mydata_complete$employ_status <- factor(mydata_complete$employ_status)
mydata_complete$marital_status <- factor(mydata_complete$marital_status)
mydata_complete$current_housing <- factor(mydata_complete$current_housing)
mydata_complete$current_zip <- factor(mydata_complete$current_zip)
#mydata_complete$have_health_insurance <- factor(mydata_complete$have_health_insurance)
mydata_complete$access_vehicle <- factor(mydata_complete$access_vehicle)
mydata_complete$water_source <- factor(mydata_complete$water_source)
mydata_complete$water_test_event_ind <- factor(mydata_complete$water_test_event_ind)
mydata_complete$collapse_last_test <- factor(mydata_complete$collapse_last_test)
mydata_complete$get_free_bottled_water <- factor(mydata_complete$get_free_bottled_water)
mydata_complete$used_water_pod <- factor(mydata_complete$used_water_pod)
mydata_complete$waterline_replaced_eventind <- factor(mydata_complete$waterline_replaced_eventind)
mydata_complete$water_line_status_after_april2014 <- factor(mydata_complete$water_line_status_after_april2014)
mydata_complete$water_bottle_disposal <- factor(mydata_complete$water_bottle_disposal)
mydata_complete$bottled_water_other <- factor(mydata_complete$bottled_water_other)
mydata_complete$bottled_water_cleaning <- factor(mydata_complete$bottled_water_cleaning)
mydata_complete$bottled_water_teeth <- factor(mydata_complete$bottled_water_teeth)
mydata_complete$bottled_water_bathe <- factor(mydata_complete$bottled_water_bathe)
mydata_complete$bottled_water_cook <- factor(mydata_complete$bottled_water_cook)
mydata_complete$bottled_water_drink <- factor(mydata_complete$bottled_water_drink)
mydata_complete$current_water_source <- factor(mydata_complete$current_water_source)
mydata_complete$blood_test_before_april_2014 <- factor(mydata_complete$blood_test_before_april_2014)
mydata_complete$blood_test_after_april_2014 <- factor(mydata_complete$blood_test_after_april_2014)
mydata_complete$elevated_lead_before_april_2014 <- factor(mydata_complete$elevated_lead_before_april_2014)
mydata_complete$elevated_lead_after_april_2014 <- factor(mydata_complete$elevated_lead_after_april_2014)
mydata_complete$no_free_bottles_long_lines <- factor(mydata_complete$no_free_bottles_long_lines)
mydata_complete$no_free_bottles_bad_hours <- factor(mydata_complete$no_free_bottles_bad_hours)
mydata_complete$no_free_bottles_no_transport <- factor(mydata_complete$no_free_bottles_no_transport)
mydata_complete$no_free_bottles_too_far <- factor(mydata_complete$no_free_bottles_too_far)
mydata_complete$no_free_bottles_supply_limit <- factor(mydata_complete$no_free_bottles_supply_limit)
mydata_complete$no_free_bottles_dont_know <- factor(mydata_complete$no_free_bottles_dont_know)
mydata_complete$no_free_bottles_use_tap <- factor(mydata_complete$no_free_bottles_use_tap)
mydata_complete$no_free_bottles_other <- factor(mydata_complete$no_free_bottles_other)
mydata_complete$drink_tap_in_three_years <- factor(mydata_complete$drink_tap_in_three_years, levels = c("Strongly Disagree", "Disagree", "Neutral", "Agree", "Strongly Agree"), ordered = TRUE)
mydata_complete$bottle_site <- factor(mydata_complete$bottle_site)

imp <- mice(mydata_complete, maxit = 0)

predM <- imp$predictorMatrix
meth <- imp$method

set.seed(217)
imp2 <- mice(mydata_complete, m = 25, print = FALSE)
```

Now we create derived variables:
```{r, echo=FALSE, message=FALSE, warning=FALSE, cache = TRUE}
imp_long <- mice::complete(imp2, action="long", include = TRUE)
imp_long$zip2 <- case_when(
  imp_long$current_zip=="48502" | imp_long$current_zip=="48503" ~ "48502/48503",
  imp_long$current_zip=="48504" ~ "48504",
  imp_long$current_zip=="48505" ~ "48505",
  imp_long$current_zip=="48506" ~ "48506",
  imp_long$current_zip=="48507" | imp_long$current_zip=="48532" ~ "48507/48532"
)

imp_long$educ_level2 <- case_when(
  imp_long$educ_level=="Less than a high school diploma" | imp_long$educ_level=="High school degree or equivalent (e.g. GED)" ~ "High school or less",
  imp_long$educ_level=="Some college, no degree" ~ "Some college",
  imp_long$educ_level=="Associate degree" | imp_long$educ_level=="Bachelor's degree" | imp_long$educ_level=="Master's, Doctoral or Professional degree" ~ "Associate's or more"
)

imp_long$public_benefits2 <- ifelse(imp_long$public_benefits=="Yes", "Yes", "No")
imp_long$gender2 <- ifelse(imp_long$gender=="Transgender Woman", "Female", imp_long$gender)

imp_long$marital_status2 <- case_when(
  imp_long$marital_status=="Single (never married)" ~ "Single (never married)",
  imp_long$marital_status=="Married" ~ "Married", 
  imp_long$marital_status=="Separated" | imp_long$marital_status=="Divorced" | imp_long$marital_status=="Widowed" ~ "Separated"
)

imp_long$employ_status2 <- case_when(
  imp_long$employ_status=="Jobs/sporadic" | imp_long$employ_status=="Other" ~ "Other",
  imp_long$employ_status=="Retired" ~ "Retired",
  imp_long$employ_status=="Employed" ~ "Employed",
  imp_long$employ_status=="Unemployed" ~ "Unemployed",
  is.na(imp_long$employ_status) ~ NA_character_)

imp_long$reth2 <- case_when(
  imp_long$reth=="Black" ~ "Black",
  imp_long$reth=="White" ~ "White",
  imp_long$reth=="Mixed Race" | imp_long$reth=="American Indian" | imp_long$reth=="Other race" | imp_long$reth=="Asian" | imp_long$reth=="Hispanic/Latinx" ~ "Other"
)

imp_long$current_housing2 <- case_when(
  imp_long$current_housing=="You own a home" ~ "Owned home",
  imp_long$current_housing=="You rent a home" ~ "Rented home",
  imp_long$current_housing=="You are staying with family or a friend" | imp_long$current_housing=="You are at a shelter" | imp_long$current_housing=="You are crashing/couch-surfing somewhere" | imp_long$current_housing=="You are on the streets" ~ "Unstable housing",
  imp_long$current_housing=="You are at a retirement facility" | imp_long$current_housing=="You are in an in-patient setting (e.g., rehab)" | imp_long$current_housing=="Other" ~ "Other housing"
)

imp_long$have_health_insurance2 <- ifelse(imp_long$have_health_insurance=="Unsure/Don't know", "No", imp_long$have_health_insurance)

imp_long$water_source2 <- ifelse(imp_long$water_source=="Tap water through the City of Flint (Water Department etc.)", "Tap water through the City of Flint", "Other")

imp_long$current_water_source2 <- case_when(
  imp_long$current_water_source=="Bottled water" | imp_long$current_water_source=="I don't drink any water at home" ~ "Bottled water",
  imp_long$current_water_source=="Tap water (with a filter)" | imp_long$current_water_source=="Water from a filtration system" ~ "Filtered tap water",
  imp_long$current_water_source=="Tap water (without a filter)" ~ "Unfiltered tap water"
)

imp_long$zip2 <- with(imp_long, factor(imp_long$zip2))
imp_long$educ_level2 <- with(imp_long, factor(imp_long$educ_level2, levels = c("High school or less", "Some college", "Associate's or more")))
imp_long$public_benefits2 <- with(imp_long, factor(imp_long$public_benefits2))
imp_long$marital_status2 <- with(imp_long, factor(imp_long$marital_status2))
imp_long$employ_status2 <- with(imp_long, factor(imp_long$employ_status2))
imp_long$reth2 <- with(imp_long, factor(imp_long$reth2))
imp_long$current_housing2 <- with(imp_long, factor(imp_long$current_housing2, levels = c("Owned home", "Rented home", "Unstable housing", "Other housing")))
imp_long$have_health_insurance2 <- with(imp_long, factor(imp_long$have_health_insurance2))
imp_long$water_source2 <- with(imp_long, factor(imp_long$water_source2, levels = c("Tap water through the City of Flint", "Other")))
imp_long$current_water_source2 <- with(imp_long, factor(imp_long$current_water_source2, levels = c("Unfiltered tap water", "Filtered tap water", "Bottled water"), ordered = TRUE))

var_label(imp_long$age) <- "Age"
var_label(imp_long$gender2) <- "Gender"
var_label(imp_long$reth2) <- "Race"
var_label(imp_long$educ_level2) <- "Education"
var_label(imp_long$public_benefits) <- "Public benefits status"
var_label(imp_long$employ_status2) <- "Employment status"
var_label(imp_long$marital_status2) <- "Marital status"
var_label(imp_long$years_cityflint) <- "Years lived in Flint"
var_label(imp_long$current_housing2) <- "Housing situation"
var_label(imp_long$zip2) <- "Zipcode"
var_label(imp_long$have_health_insurance2) <- "Health insurance"
var_label(imp_long$access_vehicle) <- "Vehicle access"
var_label(imp_long$water_source2) <- "Home water source"
var_label(imp_long$num_time_notice) <- "Months to notice a problem"
var_label(imp_long$watertestnum) <- "Number of water lead tests"
var_label(imp_long$water_test_surv) <- "Months from water switch to last water test"
var_label(imp_long$collapse_last_test) <- "Last water test result"
var_label(imp_long$blood_test_after_april_2014) <- "Got blood lead test after April 2014"
var_label(imp_long$blood_test_before_april_2014) <- "Got blood lead test before April 2014"
var_label(imp_long$elevated_lead_after_april_2014) <- "Diagnosed with elevated blood lead after April 2014"
var_label(imp_long$current_water_source2) <- "Current drinking water source"
var_label(imp_long$bottled_water_drink) <- "Uses bottled water to drink"
var_label(imp_long$bottled_water_cook) <- "Uses bottled water to cook"
var_label(imp_long$bottled_water_bathe) <- "Uses bottled water to bathe"
var_label(imp_long$bottled_water_teeth) <- "Uses bottled water to brush teeth"
var_label(imp_long$bottled_water_cleaning) <- "Uses bottled water to clean"
var_label(imp_long$bottled_water_other) <- "Uses bottled water for another purpose"
var_label(imp_long$num_bottles_per_day) <- "Number of standard (16.9 oz.) bottles of water used per day"
var_label(imp_long$number_children) <- "Number of children in care"
var_label(imp_long$water_bottle_disposal) <- "Disposal method for bottled water"
var_label(imp_long$water_line_status_after_april2014) <- "Current status of water line"
var_label(imp_long$water_line_surv) <- "Months from water switch to water line replacement"
var_label(imp_long$used_water_pod) <- "Ever used water pod (free bottled water)"
var_label(imp_long$get_free_bottled_water) <- "Currently uses water pod (free bottled water)"
var_label(imp_long$no_free_bottles_dont_know) <- "Didn't use water pods because didn't know about them"
var_label(imp_long$no_free_bottles_supply_limit) <- "Didn't use water pods because of supply limits"
var_label(imp_long$no_free_bottles_too_far) <- "Didn't use water pods because too far away"
var_label(imp_long$no_free_bottles_no_transport) <- "Didn't use water pods because of lack of transportation"
var_label(imp_long$no_free_bottles_bad_hours) <- "Didn't use water pods because they had bad hours"
var_label(imp_long$no_free_bottles_long_lines) <- "Didn't use water pods because of long lines"
var_label(imp_long$no_free_bottles_use_tap) <- "Didn't use water pods because I used my tap water"
var_label(imp_long$no_free_bottles_other) <- "Didn't use water pods for some other reason"
var_label(imp_long$bottle_site) <- "Water pod site"
var_label(imp_long$drink_tap_in_three_years) <- "Are you likely to drink the tap water in three years (~June 2022)?"
var_label(imp_long$confidants2) <- "Network size"
var_label(imp_long$bllpost2014_child) <- "Number of children with elevated BLLs"
var_label(imp_long$testpost2014_child) <- "Number of children tested for lead post-2014"
var_label(imp_long$bottledwater_child) <- "Number of children drinking bottled water"

imp_long_mids<-as.mids(imp_long)
```

Now we get our descriptives:
```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_predictors <- CreateTableOne(vars=c("age", "gender2", "educ_level2","marital_status2", "public_benefits", "have_health_insurance2", "employ_status2","years_cityflint","current_housing2","zip2","access_vehicle","water_source2","number_children","confidants2"), strata = "reth2", data=filter(imp_long, .imp==0), includeNA = TRUE, argsNormal = list(NULL), argsNonNormal = list(var.equal = TRUE))

test3 <- print(table_predictors, contDigits=1, varLabel = T, printToggle=FALSE)
test3["Age (mean (SD))", "p"] <- round(kruskal.test(age ~ reth2, data = filter(imp_long, .imp==0))$p.value, digits = 2)
test3["Years lived in Flint (mean (SD))", "p"] <- round(kruskal.test(years_cityflint ~ reth2, data = filter(imp_long, .imp==0))$p.value, digits = 2)
test3["Number of children in care (mean (SD))", "p"] <- round(kruskal.test(number_children ~ reth2, data = filter(imp_long, .imp==0))$p.value, digits = 2)
test3["Network size (mean (SD))", "p"] <- round(kruskal.test(confidants2 ~ reth2, data = filter(imp_long, .imp==0))$p.value, digits = 2)

table_outcomes <- CreateTableOne(vars=c("num_time_notice","watertestnum","water_test_surv", "collapse_last_test", "blood_test_after_april_2014","elevated_lead_after_april_2014","current_water_source2","used_water_pod","get_free_bottled_water","bottle_site","bottled_water_drink","bottled_water_cook","bottled_water_bathe","bottled_water_teeth","bottled_water_cleaning","bottled_water_other", "num_bottles_per_day","water_bottle_disposal","no_free_bottles_dont_know","no_free_bottles_supply_limit","no_free_bottles_too_far","no_free_bottles_no_transport","no_free_bottles_bad_hours","no_free_bottles_long_lines","no_free_bottles_use_tap","no_free_bottles_other","water_line_status_after_april2014","water_line_surv", "drink_tap_in_three_years"), strata = "reth2", data=filter(imp_long, .imp==0), includeNA = TRUE, argsNormal = list(NULL), argsNonNormal = list(var.equal = TRUE))

table_predictors_notstrat <- CreateTableOne(vars=c("age", "gender2", "educ_level2","marital_status2", "public_benefits", "have_health_insurance2", "employ_status2","years_cityflint","current_housing2","zip2","access_vehicle","water_source2","number_children","confidants2"), data=filter(imp_long, .imp==0), argsNormal = list(NULL), argsNonNormal = list(var.equal = TRUE), includeNA=TRUE)

table_outcomes_notstrat <- CreateTableOne(vars=c("num_time_notice","watertestnum","water_test_surv", "collapse_last_test", "blood_test_after_april_2014","elevated_lead_after_april_2014","current_water_source2","used_water_pod","get_free_bottled_water","bottle_site","bottled_water_drink","bottled_water_cook","bottled_water_bathe","bottled_water_teeth","bottled_water_cleaning","bottled_water_other", "num_bottles_per_day","water_bottle_disposal","no_free_bottles_dont_know","no_free_bottles_supply_limit","no_free_bottles_too_far","no_free_bottles_no_transport","no_free_bottles_bad_hours","no_free_bottles_long_lines","no_free_bottles_use_tap","no_free_bottles_other","water_line_status_after_april2014","water_line_surv", "drink_tap_in_three_years"), data=filter(imp_long, .imp==0), argsNormal = list(NULL), argsNonNormal = list(var.equal = TRUE), includeNA=TRUE)

test <- print(table_outcomes, contDigits=1, varLabel = T, printToggle=FALSE)
test2 <- print(table_outcomes_notstrat, contDigits=1, varLabel = T, printToggle=FALSE)

test["Months to notice a problem (mean (SD))", "p"] <- round(kruskal.test(num_time_notice ~ reth2, data = filter(imp_long, .imp==0))$p.value, digits = 2)
test["Number of water lead tests (mean (SD))", "p"] <- round(kruskal.test(watertestnum ~ reth2, data = filter(imp_long, .imp==0))$p.value, digits = 2)

km_waterline <- survfit(Surv(water_line_surv, as.numeric(waterline_replaced_eventind)-1) ~ 1, data = filter(imp_long, .imp==0, waterline_cured_ind==0))
print(km_waterline)
km_waterline_reth <- survfit(Surv(water_line_surv, as.numeric(waterline_replaced_eventind)-1) ~ reth2, data = filter(imp_long, .imp==0, waterline_cured_ind==0), conf.type = "log-log")
print(km_waterline_reth)

waterline_reth <- c("55 (53, 58)", "53 (46, 64)", "54 (49, 56)")
waterline_overall <- "54 (53, 56)"

testresults <- survdiff(Surv(water_line_surv, as.numeric(waterline_replaced_eventind)-1) ~ reth2, data = filter(imp_long, .imp==0, waterline_cured_ind==0))

test["Months from water switch to water line replacement (mean (SD))", c("Black", "Other", "White")] <- waterline_reth
test2["Months from water switch to water line replacement (mean (SD))", "Overall"] <- waterline_overall
test["Months from water switch to water line replacement (mean (SD))", "p"] <- round(dchisq(testresults$chisq, df = 2)*2, digits = 2)

km_watertest <- survfit(Surv(water_test_surv, as.numeric(water_test_event_ind)-1) ~ 1, data = filter(imp_long, .imp==0), conf.type = "log-log")
print(km_watertest)
km_watertest_reth <- survfit(Surv(water_test_surv, as.numeric(water_test_event_ind)-1) ~ reth2, data = filter(imp_long, .imp==0), conf.type = "log-log")
print(km_watertest_reth)

watertest_reth <- c("53 (53, 58)", "62 (57, 64)", "55 (53, 62)")
watertest_overall <- "56 (53, 59)"

test["Months from water switch to last water test (mean (SD))", c("Black", "Other", "White")] <- watertest_reth
test2["Months from water switch to last water test (mean (SD))", "Overall"] <- watertest_overall
test["Months from water switch to last water test (mean (SD))", "p"] <- round(dchisq(survdiff(Surv(water_test_surv, as.numeric(water_test_event_ind)-1) ~ reth2, data = filter(imp_long, .imp==0))$chisq, df = 2)*2, digits = 2)

subdat <- filter(imp_long, .imp==0)
medians <- c(median(subdat$num_time_notice[subdat$reth2=="Black"], na.rm = TRUE), median(subdat$num_time_notice[subdat$reth2=="Other"], na.rm = TRUE), median(subdat$num_time_notice[subdat$reth2=="White"], na.rm = TRUE))
mins <- c(min(subdat$num_time_notice[subdat$reth2=="Black"], na.rm = TRUE), min(subdat$num_time_notice[subdat$reth2=="Other"], na.rm = TRUE), min(subdat$num_time_notice[subdat$reth2=="White"], na.rm = TRUE))
maxs <- c(max(subdat$num_time_notice[subdat$reth2=="Black"], na.rm = TRUE), max(subdat$num_time_notice[subdat$reth2=="Other"], na.rm = TRUE), max(subdat$num_time_notice[subdat$reth2=="White"], na.rm = TRUE))

test["Months to notice a problem (mean (SD))", c("Black", "Other", "White")] <- paste0(round(medians, digits = 1), " (", round(mins, digits = 1), ", ", round(maxs, digits = 1), ")")

test2["Months to notice a problem (mean (SD))", "Overall"] <- paste0(round(median(subdat$num_time_notice, na.rm = TRUE), digits = 1), " (", round(min(subdat$num_time_notice, na.rm = TRUE), digits = 1), ", ", round(max(subdat$num_time_notice, na.rm = TRUE), digits = 1), ")")

kable(test3, booktabs=T) %>% add_indent(c(5:7, 9:11, 13:15, 18:22, 25:28, 30:34, 37:39)) %>% kable_styling(position="center")

kable(print(table_predictors_notstrat, contDigits=1, varLabel = T, printToggle=FALSE), booktabs=T) %>% add_indent(c(5:7, 9:11, 13:15, 18:22, 25:28, 30:34, 37:39)) %>% kable_styling(position="center")

kable(test, booktabs=T) %>% add_indent(c(6:10, 12:15, 17:20, 22:25, 28:30, 32:47, 56:60, 70:77, 80:85)) %>% kable_styling(position="center")

kable(test2, booktabs=T) %>% add_indent(c(6:10, 12:15, 17:20, 22:25, 28:30, 32:47, 56:60, 70:77, 80:85)) %>% kable_styling(position="center")
```

We generate some figures:
```{r, echo=FALSE, message=FALSE, warning=FALSE, eval = TRUE}
imp_long$watertestbinary <- as.numeric(imp_long$watertestnum > 0)
imp_long$watertestbinary[is.na(imp_long$watertestbinary)] <- 2
imp_long$watertestbinary <- factor(imp_long$watertestbinary, levels=c(1, 0, 2), labels=c("Yes", "No", "Missing"), ordered = TRUE)
imp_long$collapse_last_testmis <- case_when(
  imp_long$collapse_last_test=="Acceptable levels of lead" ~ "Safe",
  imp_long$collapse_last_test=="Unsure if tested" | is.na(imp_long$collapse_last_test)~ "Missing",
  imp_long$collapse_last_test=="Not tested" ~ "Not \ntested",
  imp_long$collapse_last_test=="Unsafe levels of lead" ~ "Unsafe"
)

imp_long$collapse_last_testmis <- factor(imp_long$collapse_last_testmis, levels=c("Safe", "Unsafe", "Not \ntested", "Missing"), ordered = TRUE)

imp_long$blood_test_after_april_2014mis <- case_when(
  imp_long$blood_test_after_april_2014=="Yes" ~ "Yes",
  imp_long$blood_test_after_april_2014=="No" ~ "No",
  imp_long$blood_test_after_april_2014=="Unsure" | is.na(imp_long$blood_test_after_april_2014) ~ "Missing"
)

imp_long$elevated_lead_after_april_2014mis <- case_when(
  imp_long$elevated_lead_after_april_2014=="Yes" ~ "Yes",
  imp_long$elevated_lead_after_april_2014=="No" ~ "No",
  imp_long$elevated_lead_after_april_2014=="Unsure" | is.na(imp_long$elevated_lead_after_april_2014) ~ "Missing"
)

alluvial_dat1 <- filter(imp_long, .imp==0) %>% dplyr::select(watertestbinary, collapse_last_testmis, blood_test_after_april_2014mis, elevated_lead_after_april_2014mis) %>% count(watertestbinary, collapse_last_testmis, blood_test_after_april_2014mis, elevated_lead_after_april_2014mis) %>% rename(Freq = n)

ggplot(alluvial_dat1, aes(y = Freq, axis1 = watertestbinary, axis2 = blood_test_after_april_2014mis, axis3 = elevated_lead_after_april_2014mis, axis4 = collapse_last_testmis)) + geom_alluvium(aes(fill = watertestbinary)) + geom_stratum(width = 1/3) + geom_text(stat = "stratum", aes(label = after_stat(stratum))) + scale_x_continuous(breaks = 1:4, labels = c("Had water test", "Had blood test", "Elevated BLLs", "Last water test result")) + theme_bw() + theme(legend.position = "none") + scale_fill_manual(values = c("mediumpurple4", "darkred", "gray74")) + ylab("Number of Respondents") + ggtitle("Participation in Testing")

alluvial_dat2 <- filter(imp_long, .imp==0, bottled_water_drink=="Yes") %>% dplyr::select(bottled_water_cook, bottled_water_bathe, bottled_water_teeth, bottled_water_cleaning, bottled_water_other) %>% count(bottled_water_cook, bottled_water_bathe, bottled_water_teeth, bottled_water_cleaning, bottled_water_other) %>% rename(Freq = n)

ggplot(alluvial_dat2, aes(y = Freq, axis1 = bottled_water_cook, axis2 = bottled_water_bathe, axis3 = bottled_water_teeth, axis4 = bottled_water_cleaning, axis5 = bottled_water_other)) + geom_alluvium(aes(fill = bottled_water_cook)) + geom_stratum(width = 1/8) + geom_text(stat = "stratum", aes(label = after_stat(stratum))) + scale_x_continuous(breaks = 1:5, labels = c("Cooking", "Bathing", "Toothbrushing", "Cleaning", "Other use")) + theme_bw() + theme(legend.position = "none") + scale_fill_manual(values = c("mediumpurple4", "darkred")) + ylab("Number of Respondents") + ggtitle("Bottled Water Uses in Addition to Drinking")

alluvial_dat3 <- filter(imp_long, .imp==0, get_free_bottled_water=="No") %>% dplyr::select(no_free_bottles_dont_know, no_free_bottles_supply_limit, no_free_bottles_too_far, no_free_bottles_no_transport, no_free_bottles_bad_hours, no_free_bottles_long_lines, no_free_bottles_use_tap, no_free_bottles_other) %>% count(no_free_bottles_dont_know, no_free_bottles_supply_limit, no_free_bottles_too_far, no_free_bottles_no_transport, no_free_bottles_bad_hours, no_free_bottles_long_lines, no_free_bottles_use_tap, no_free_bottles_other) %>% rename(Freq = n)

ggplot(alluvial_dat3, aes(y = Freq, axis1 = no_free_bottles_dont_know, axis2 = no_free_bottles_supply_limit, axis3 = no_free_bottles_too_far, axis4 = no_free_bottles_no_transport, axis5 = no_free_bottles_bad_hours, axis6 = no_free_bottles_long_lines, axis7 = no_free_bottles_use_tap, axis8 = no_free_bottles_other)) + geom_alluvium(aes(fill = no_free_bottles_dont_know)) + geom_stratum(width = 1/8) + geom_text(stat = "stratum", aes(label = after_stat(stratum))) + scale_x_continuous(breaks = 1:8, labels = c("Didn't know", "Supply limit", "Too far", "No transportation", "Bad hours", "Long lines", "Use tap instead", "Other reason")) + theme_bw() + theme(legend.position = "none", axis.text.x = element_text(angle=45, vjust = 0.5)) + scale_fill_manual(values = c("mediumpurple4", "darkred")) + ylab("Number of Respondents") + ggtitle("Reasons for Not Currently Using Water Pods")

timingdat <- data.frame("Date" = c(mydata$notice_date, mydata$last_test_date, mydata$waterline_date), "Type" = rep(c("Noticed water issue", "Last water test", "Waterline replaced"), each = nrow(mydata)))

timingdat$Type <- factor(timingdat$Type, levels=c("Noticed water issue", "Last water test", "Waterline replaced"), ordered = TRUE)

ggplot(timingdat, aes(x = Date, fill = Type, color = Type)) + geom_histogram(position = "dodge") + ylab("Number of Respondents") + theme_bw() + scale_fill_manual(values = c("mediumpurple4", "gray74", "darkred")) + scale_color_manual(values = c("mediumpurple4", "gray74", "darkred")) + ggtitle("Timing of Water Crisis Response") #+ coord_cartesian(ylim = c(0, 60))
```

We do the modeling:
```{r, echo=FALSE, message=FALSE, warning=FALSE}
effect_estims <- data.frame("Outcome" = NA, "Predictor" = NA, "Estimate" = NA, "SE" = NA, "Pval" = NA)

time_to_notice <- with(imp_long_mids, lm(log(num_time_notice) ~ age + reth2 + gender2 + educ_level2 + marital_status2 + public_benefits + have_health_insurance2 + employ_status2 + years_cityflint + current_housing2 + water_source2 + access_vehicle + number_children + confidants2))

summary(pool(time_to_notice))

subdat <- data.frame("Outcome" = rep("Time to notice", 22), "Predictor" = c("Age", "Other race (ref: Black race)", "White race (ref: Black race)", "Male (ref: Female)", "Some college (ref: High school or less)", "Associate's degree or more (ref: High school or less", "Separated (ref: Married)", "Single (ref: Married)", "No public benefits (ref: Receives public benefits)", "Unsure if receives public benefits (ref: Receives public benefits)", "Health insured (ref: Uninsured)", "Other employment (ref: Employed)", "Retired (ref: Employed)", "Unemployed (ref: Employed)", "Years lived in Flint", "Rented home (ref: Owned home)", "Unstable housing (ref: Owned home)", "Other housing (ref: Owned home)", "Other water source (ref: Flint tap water)", "Has vehicle (ref: No vehicle)", "Number of children", "Network size"), "Estimate" = summary(pool(time_to_notice))$estimate[-1], "SE" = summary(pool(time_to_notice))$std.error[-1], "Pval" = summary(pool(time_to_notice))$p.value[-1])

effect_estims <- rbind(effect_estims, subdat)

number_water_tests <- with(imp_long_mids, lm(log(watertestnum + 0.1) ~ age + reth2 + gender2 + educ_level2 + marital_status2 + public_benefits + have_health_insurance2 + employ_status2 + years_cityflint + current_housing2 + access_vehicle + water_source2 + number_children + confidants2))

summary(pool(number_water_tests))

subdat <- data.frame("Outcome" = rep("Number of water tests", 22), "Predictor" = c("Age", "Other race (ref: Black race)", "White race (ref: Black race)", "Male (ref: Female)", "Some college (ref: High school or less)", "Associate's degree or more (ref: High school or less", "Separated (ref: Married)", "Single (ref: Married)", "No public benefits (ref: Receives public benefits)", "Unsure if receives public benefits (ref: Receives public benefits)", "Health insured (ref: Uninsured)", "Other employment (ref: Employed)", "Retired (ref: Employed)", "Unemployed (ref: Employed)", "Years lived in Flint", "Rented home (ref: Owned home)", "Unstable housing (ref: Owned home)", "Other housing (ref: Owned home)", "Other water source (ref: Flint tap water)", "Has vehicle (ref: No vehicle)", "Number of children", "Network size"), "Estimate" = summary(pool(number_water_tests))$estimate[-1], "SE" = summary(pool(number_water_tests))$std.error[-1], "Pval" = summary(pool(number_water_tests))$p.value[-1])

effect_estims <- rbind(effect_estims, subdat)

last_water_test_time <- with(imp_long_mids, coxph(Surv(water_test_surv, as.numeric(water_test_event_ind)-1) ~ age + reth2 + gender2 + educ_level2 + marital_status2 + public_benefits + have_health_insurance2 + employ_status2 + years_cityflint + current_housing2 + access_vehicle + water_source2 + number_children + confidants2))

summary(pool(last_water_test_time))

subdat <- data.frame("Outcome" = rep("Time to last water test", 22), "Predictor" = c("Age", "Other race (ref: Black race)", "White race (ref: Black race)", "Male (ref: Female)", "Some college (ref: High school or less)", "Associate's degree or more (ref: High school or less", "Separated (ref: Married)", "Single (ref: Married)", "No public benefits (ref: Receives public benefits)", "Unsure if receives public benefits (ref: Receives public benefits)", "Health insured (ref: Uninsured)", "Other employment (ref: Employed)", "Retired (ref: Employed)", "Unemployed (ref: Employed)", "Years lived in Flint", "Rented home (ref: Owned home)", "Unstable housing (ref: Owned home)", "Other housing (ref: Owned home)", "Other water source (ref: Flint tap water)", "Has vehicle (ref: No vehicle)", "Number of children", "Network size"), "Estimate" = summary(pool(last_water_test_time))$estimate, "SE" = summary(pool(last_water_test_time))$std.error, "Pval" = summary(pool(last_water_test_time))$p.value)

effect_estims <- rbind(effect_estims, subdat)

last_test_result <- with(imp_long_mids, glm(as.numeric(collapse_last_test=="Acceptable levels of lead") ~ age + reth2 + gender2 + educ_level2 + marital_status2 + public_benefits + have_health_insurance2 + employ_status2 + years_cityflint + current_housing2 + access_vehicle + water_source2 + number_children + confidants2, family = "binomial"))

summary(pool(last_test_result))

subdat <- data.frame("Outcome" = rep("Last water test result", 22), "Predictor" = c("Age", "Other race (ref: Black race)", "White race (ref: Black race)", "Male (ref: Female)", "Some college (ref: High school or less)", "Associate's degree or more (ref: High school or less", "Separated (ref: Married)", "Single (ref: Married)", "No public benefits (ref: Receives public benefits)", "Unsure if receives public benefits (ref: Receives public benefits)", "Health insured (ref: Uninsured)", "Other employment (ref: Employed)", "Retired (ref: Employed)", "Unemployed (ref: Employed)", "Years lived in Flint", "Rented home (ref: Owned home)", "Unstable housing (ref: Owned home)", "Other housing (ref: Owned home)", "Other water source (ref: Flint tap water)", "Has vehicle (ref: No vehicle)", "Number of children", "Network size"), "Estimate" = summary(pool(last_test_result))$estimate[-1], "SE" = summary(pool(last_test_result))$std.error[-1], "Pval" = summary(pool(last_test_result))$p.value[-1])

effect_estims <- rbind(effect_estims, subdat)

blood_test <- with(imp_long_mids, glm(as.numeric(blood_test_after_april_2014=="Yes") ~ age + reth2 + gender2 + educ_level2 + marital_status2 + public_benefits + have_health_insurance2 + employ_status2 + years_cityflint + current_housing2 + access_vehicle + water_source2 + number_children + confidants2, family = "binomial"))

summary(pool(blood_test))

subdat <- data.frame("Outcome" = rep("Blood test", 22), "Predictor" = c("Age", "Other race (ref: Black race)", "White race (ref: Black race)", "Male (ref: Female)", "Some college (ref: High school or less)", "Associate's degree or more (ref: High school or less", "Separated (ref: Married)", "Single (ref: Married)", "No public benefits (ref: Receives public benefits)", "Unsure if receives public benefits (ref: Receives public benefits)", "Health insured (ref: Uninsured)", "Other employment (ref: Employed)", "Retired (ref: Employed)", "Unemployed (ref: Employed)", "Years lived in Flint", "Rented home (ref: Owned home)", "Unstable housing (ref: Owned home)", "Other housing (ref: Owned home)", "Other water source (ref: Flint tap water)", "Has vehicle (ref: No vehicle)", "Number of children", "Network size"), "Estimate" = summary(pool(blood_test))$estimate[-1], "SE" = summary(pool(blood_test))$std.error[-1], "Pval" = summary(pool(blood_test))$p.value[-1])

effect_estims <- rbind(effect_estims, subdat)

## This model could not be fit because of small sample sizes.
#elev_blls <- with(imp_long_mids, glm(as.numeric(elevated_lead_after_april_2014=="Yes") ~ age + reth2 + gender2 + educ_level2 + marital_status2 + public_benefits + have_health_insurance2 + employ_status2 + years_cityflint + water_source2 + number_children + confidants2, family = "binomial"))

#summary(pool(elev_blls))

#subdat <- data.frame("Outcome" = rep("Elevated BLLs", 18), "Predictor" = c("Age", "Other race (ref: Black race)", "White race (ref: Black race)", "Male (ref: Female)", "Some college (ref: High school or less)", "Associate's degree or more (ref: High school or less", "Separated (ref: Married)", "Single (ref: Married)", "No public benefits (ref: Receives public benefits)", "Unsure if receives public benefits (ref: Receives public benefits)", "Health insured (ref: Uninsured)", "Other employment (ref: Employed)", "Retired (ref: Employed)", "Unemployed (ref: Employed)", "Years lived in Flint", "Other water source (ref: Flint tap water)", "Number of children", "Network size"), "Estimate" = summary(pool(elev_blls))$estimate[-1], "SE" = summary(pool(elev_blls))$std.error[-1], "Pval" = summary(pool(elev_blls))$p.value[-1])

#effect_estims <- rbind(effect_estims, subdat)

currentwatersource <- with(imp_long_mids, polr(current_water_source2 ~ age + reth2 + gender2 + educ_level2 + marital_status2 + public_benefits + have_health_insurance2 + employ_status2 + years_cityflint + current_housing2 + access_vehicle + water_source2 + number_children + confidants2, Hess = TRUE))

summary(pool(currentwatersource))

subdat <- data.frame("Outcome" = rep("Current drinking water source", 22), "Predictor" = c("Age", "Other race (ref: Black race)", "White race (ref: Black race)", "Male (ref: Female)", "Some college (ref: High school or less)", "Associate's degree or more (ref: High school or less", "Separated (ref: Married)", "Single (ref: Married)", "No public benefits (ref: Receives public benefits)", "Unsure if receives public benefits (ref: Receives public benefits)", "Health insured (ref: Uninsured)", "Other employment (ref: Employed)", "Retired (ref: Employed)", "Unemployed (ref: Employed)", "Years lived in Flint", "Rented home (ref: Owned home)", "Unstable housing (ref: Owned home)", "Other housing (ref: Owned home)", "Other water source (ref: Flint tap water)", "Has vehicle (ref: No vehicle)", "Number of children", "Network size"), "Estimate" = summary(pool(currentwatersource))$estimate[1:22], "SE" = summary(pool(currentwatersource))$std.error[1:22], "Pval" = NA)

effect_estims <- rbind(effect_estims, subdat)

water_pod_past <- with(imp_long_mids, glm(used_water_pod ~ age + reth2 + gender2 + educ_level2 + marital_status2 + public_benefits + have_health_insurance2 + employ_status2 + years_cityflint + current_housing2 + access_vehicle + water_source2 + number_children + confidants2, family = "binomial"))

summary(pool(water_pod_past))

subdat <- data.frame("Outcome" = rep("Used water pod in the past", 22), "Predictor" = c("Age", "Other race (ref: Black race)", "White race (ref: Black race)", "Male (ref: Female)", "Some college (ref: High school or less)", "Associate's degree or more (ref: High school or less", "Separated (ref: Married)", "Single (ref: Married)", "No public benefits (ref: Receives public benefits)", "Unsure if receives public benefits (ref: Receives public benefits)", "Health insured (ref: Uninsured)", "Other employment (ref: Employed)", "Retired (ref: Employed)", "Unemployed (ref: Employed)", "Years lived in Flint", "Rented home (ref: Owned home)", "Unstable housing (ref: Owned home)", "Other housing (ref: Owned home)", "Other water source (ref: Flint tap water)", "Has vehicle (ref: No vehicle)", "Number of children", "Network size"), "Estimate" = summary(pool(water_pod_past))$estimate[-1], "SE" = summary(pool(water_pod_past))$std.error[-1], "Pval" = summary(pool(water_pod_past))$p.value[-1])

effect_estims <- rbind(effect_estims, subdat)

water_pod_now <- with(imp_long_mids, glm(get_free_bottled_water ~ age + reth2 + gender2 + educ_level2 + marital_status2 + public_benefits + have_health_insurance2 + employ_status2 + years_cityflint + current_housing2 + access_vehicle + water_source2 + number_children + confidants2, family = "binomial"))

summary(pool(water_pod_now))

subdat <- data.frame("Outcome" = rep("Use water pod now", 22), "Predictor" = c("Age", "Other race (ref: Black race)", "White race (ref: Black race)", "Male (ref: Female)", "Some college (ref: High school or less)", "Associate's degree or more (ref: High school or less", "Separated (ref: Married)", "Single (ref: Married)", "No public benefits (ref: Receives public benefits)", "Unsure if receives public benefits (ref: Receives public benefits)", "Health insured (ref: Uninsured)", "Other employment (ref: Employed)", "Retired (ref: Employed)", "Unemployed (ref: Employed)", "Years lived in Flint", "Rented home (ref: Owned home)", "Unstable housing (ref: Owned home)", "Other housing (ref: Owned home)", "Other water source (ref: Flint tap water)", "Has vehicle (ref: No vehicle)", "Number of children", "Network size"), "Estimate" = summary(pool(water_pod_now))$estimate[-1], "SE" = summary(pool(water_pod_now))$std.error[-1], "Pval" = summary(pool(water_pod_now))$p.value[-1])

effect_estims <- rbind(effect_estims, subdat)

number_bottles <- with(imp_long_mids, glm(num_bottles_per_day ~ age + reth2 + gender2 + educ_level2 + marital_status2 + public_benefits + have_health_insurance2 + employ_status2 + years_cityflint + current_housing2 + access_vehicle + water_source2 + number_children + confidants2, family = "quasipoisson"))

summary(pool(number_bottles))

subdat <- data.frame("Outcome" = rep("Number of bottles per day", 22), "Predictor" = c("Age", "Other race (ref: Black race)", "White race (ref: Black race)", "Male (ref: Female)", "Some college (ref: High school or less)", "Associate's degree or more (ref: High school or less", "Separated (ref: Married)", "Single (ref: Married)", "No public benefits (ref: Receives public benefits)", "Unsure if receives public benefits (ref: Receives public benefits)", "Health insured (ref: Uninsured)", "Other employment (ref: Employed)", "Retired (ref: Employed)", "Unemployed (ref: Employed)", "Years lived in Flint", "Rented home (ref: Owned home)", "Unstable housing (ref: Owned home)", "Other housing (ref: Owned home)", "Other water source (ref: Flint tap water)", "Has vehicle (ref: No vehicle)", "Number of children", "Network size"), "Estimate" = summary(pool(number_bottles))$estimate[-1], "SE" = summary(pool(number_bottles))$std.error[-1], "Pval" = summary(pool(number_bottles))$p.value[-1])

effect_estims <- rbind(effect_estims, subdat)

recycle_bottles <- with(imp_long_mids, glm(as.numeric(water_bottle_disposal=="I don't ever have plastic water bottles at home" | water_bottle_disposal=="Keep/save them for re-use or another purpose" | water_bottle_disposal=="Recycle them") ~ age + reth2 + gender2 + educ_level2 + marital_status2 + public_benefits + have_health_insurance2 + employ_status2 + years_cityflint + current_housing2 + access_vehicle + water_source2 + number_children + confidants2, family = "binomial"))

summary(pool(recycle_bottles))

subdat <- data.frame("Outcome" = rep("Bottle disposal", 22), "Predictor" = c("Age", "Other race (ref: Black race)", "White race (ref: Black race)", "Male (ref: Female)", "Some college (ref: High school or less)", "Associate's degree or more (ref: High school or less", "Separated (ref: Married)", "Single (ref: Married)", "No public benefits (ref: Receives public benefits)", "Unsure if receives public benefits (ref: Receives public benefits)", "Health insured (ref: Uninsured)", "Other employment (ref: Employed)", "Retired (ref: Employed)", "Unemployed (ref: Employed)", "Years lived in Flint", "Rented home (ref: Owned home)", "Unstable housing (ref: Owned home)", "Other housing (ref: Owned home)", "Other water source (ref: Flint tap water)", "Has vehicle (ref: No vehicle)", "Number of children", "Network size"), "Estimate" = summary(pool(recycle_bottles))$estimate[-1], "SE" = summary(pool(recycle_bottles))$std.error[-1], "Pval" = summary(pool(recycle_bottles))$p.value[-1])

effect_estims <- rbind(effect_estims, subdat)

waterlinestatus <- with(imp_long_mids, glm(as.numeric(water_line_status_after_april2014=="A mailing or official said that my lead/water line did not need to be replaced" | water_line_status_after_april2014=="I was offered a lead/water line replacement but I refused it" | water_line_status_after_april2014=="My lead/water lines were replaced") ~ age + reth2 + gender2 + educ_level2 + marital_status2 + public_benefits + have_health_insurance2 + employ_status2 + years_cityflint + current_housing2 + access_vehicle + water_source2 + number_children + confidants2, family = "binomial"))

summary(pool(waterlinestatus))

subdat <- data.frame("Outcome" = rep("Water line status", 22), "Predictor" = c("Age", "Other race (ref: Black race)", "White race (ref: Black race)", "Male (ref: Female)", "Some college (ref: High school or less)", "Associate's degree or more (ref: High school or less", "Separated (ref: Married)", "Single (ref: Married)", "No public benefits (ref: Receives public benefits)", "Unsure if receives public benefits (ref: Receives public benefits)", "Health insured (ref: Uninsured)", "Other employment (ref: Employed)", "Retired (ref: Employed)", "Unemployed (ref: Employed)", "Years lived in Flint", "Rented home (ref: Owned home)", "Unstable housing (ref: Owned home)", "Other housing (ref: Owned home)", "Other water source (ref: Flint tap water)", "Has vehicle (ref: No vehicle)", "Number of children", "Network size"), "Estimate" = summary(pool(waterlinestatus))$estimate[-1], "SE" = summary(pool(waterlinestatus))$std.error[-1], "Pval" = summary(pool(waterlinestatus))$p.value[-1])

effect_estims <- rbind(effect_estims, subdat)

water_line_time <- with(imp_long_mids, coxph(Surv(water_line_surv, as.numeric(waterline_replaced_eventind)-1) ~ age + reth2 + gender2 + educ_level2 + marital_status2 + public_benefits + have_health_insurance2 + employ_status2 + years_cityflint + current_housing2 + access_vehicle + water_source2 + number_children + confidants2))

summary(pool(water_line_time))

subdat <- data.frame("Outcome" = rep("Time to water line replacement", 22), "Predictor" = c("Age", "Other race (ref: Black race)", "White race (ref: Black race)", "Male (ref: Female)", "Some college (ref: High school or less)", "Associate's degree or more (ref: High school or less", "Separated (ref: Married)", "Single (ref: Married)", "No public benefits (ref: Receives public benefits)", "Unsure if receives public benefits (ref: Receives public benefits)", "Health insured (ref: Uninsured)", "Other employment (ref: Employed)", "Retired (ref: Employed)", "Unemployed (ref: Employed)", "Years lived in Flint", "Rented home (ref: Owned home)", "Unstable housing (ref: Owned home)", "Other housing (ref: Owned home)", "Other water source (ref: Flint tap water)", "Has vehicle (ref: No vehicle)", "Number of children", "Network size"), "Estimate" = summary(pool(water_line_time))$estimate, "SE" = summary(pool(water_line_time))$std.error, "Pval" = summary(pool(water_line_time))$p.value)

effect_estims <- rbind(effect_estims, subdat)

drinktap <- with(imp_long_mids, polr(drink_tap_in_three_years ~ age + reth2 + gender2 + educ_level2 + marital_status2 + public_benefits + have_health_insurance2 + employ_status2 + years_cityflint + current_housing2 + access_vehicle + water_source2 + number_children + confidants2, Hess = TRUE))

summary(pool(drinktap))

subdat <- data.frame("Outcome" = rep("Likely to drink tap in three years?", 22), "Predictor" = c("Age", "Other race (ref: Black race)", "White race (ref: Black race)", "Male (ref: Female)", "Some college (ref: High school or less)", "Associate's degree or more (ref: High school or less", "Separated (ref: Married)", "Single (ref: Married)", "No public benefits (ref: Receives public benefits)", "Unsure if receives public benefits (ref: Receives public benefits)", "Health insured (ref: Uninsured)", "Other employment (ref: Employed)", "Retired (ref: Employed)", "Unemployed (ref: Employed)", "Years lived in Flint", "Rented home (ref: Owned home)", "Unstable housing (ref: Owned home)", "Other housing (ref: Owned home)", "Other water source (ref: Flint tap water)", "Has vehicle (ref: No vehicle)", "Number of children", "Network size"), "Estimate" = summary(pool(drinktap))$estimate[1:22], "SE" = summary(pool(drinktap))$std.error[1:22], "Pval" = NA)

effect_estims <- rbind(effect_estims, subdat)

#child_test <- with(imp_long_mids, glm(testpost2014_child ~ age + reth2 + gender2 + educ_level2 + marital_status2 + public_benefits + have_health_insurance2 + employ_status2 + years_cityflint + current_housing2 + access_vehicle + water_source2 + confidants2 + factor(blood_test_after_april_2014, ordered = FALSE) + offset(number_children), family = "quasipoisson"))

#summary(pool(child_test))

#subdat <- data.frame("Outcome" = rep("Got child tested", 23), "Predictor" = c("Age", "Other race (ref: Black race)", "White race (ref: Black race)", "Male (ref: Female)", "Some college (ref: High school or less)", "Associate's degree or more (ref: High school or less", "Separated (ref: Married)", "Single (ref: Married)", "No public benefits (ref: Receives public benefits)", "Unsure if receives public benefits (ref: Receives public benefits)", "Health insured (ref: Uninsured)", "Other employment (ref: Employed)", "Retired (ref: Employed)", "Unemployed (ref: Employed)", "Years lived in Flint", "Rented home (ref: Owned home)", "Unstable housing (ref: Owned home)", "Other housing (ref: Owned home)", "Other water source (ref: Flint tap water)", "Has vehicle (ref: No vehicle)", "Network size", "Adult not tested (ref: Adult tested)", "Adult unsure if tested (ref: Adult tested)"), "Estimate" = summary(pool(child_test))$estimate[-1], "SE" = summary(pool(child_test))$std.error[-1], "Pval" = summary(pool(child_test))$p.value[-1])

#effect_estims <- rbind(effect_estims, subdat)

##This model not included because of stability issues:

#child_bll <- with(imp_long_mids, glm(bllpost2014_child ~ age + reth2 + gender2 + educ_level2 + marital_status2 + public_benefits + have_health_insurance2 + employ_status2 + years_cityflint + current_housing2 + access_vehicle + water_source2 + confidants2 + as.numeric(elevated_lead_after_april_2014=="Yes") + offset(number_children), family = "quasipoisson"))

#summary(pool(child_bll))

#subdat <- data.frame("Outcome" = rep("Child had elevated BLLs", 22), "Predictor" = c("Age", "Other race (ref: Black race)", "White race (ref: Black race)", "Male (ref: Female)", "Some college (ref: High school or less)", "Associate's degree or more (ref: High school or less", "Separated (ref: Married)", "Single (ref: Married)", "No public benefits (ref: Receives public benefits)", "Unsure if receives public benefits (ref: Receives public benefits)", "Health insured (ref: Uninsured)", "Other employment (ref: Employed)", "Retired (ref: Employed)", "Unemployed (ref: Employed)", "Years lived in Flint", "Rented home (ref: Owned home)", "Unstable housing (ref: Owned home)", "Other housing (ref: Owned home)", "Other water source (ref: Flint tap water)", "Has vehicle (ref: No vehicle)", "Network size", "Adult has elevated BLLs (ref: Adult has normal/unsure BLLs)"), "Estimate" = summary(pool(child_bll))$estimate[-1], "SE" = summary(pool(child_bll))$std.error[-1], "Pval" = summary(pool(child_bll))$p.value[-1])

#effect_estims <- rbind(effect_estims, subdat)

#child_bottled <- with(imp_long_mids, glm(bottledwater_child ~ age + reth2 + gender2 + educ_level2 + marital_status2 + public_benefits + have_health_insurance2 + employ_status2 + years_cityflint + current_housing2 + access_vehicle + water_source2 + confidants2 + factor(current_water_source2, ordered = FALSE) + offset(number_children), family = "quasipoisson"))

#summary(pool(child_bottled))

#subdat <- data.frame("Outcome" = rep("Child uses bottled water", 23), "Predictor" = c("Age", "Other race (ref: Black race)", "White race (ref: Black race)", "Male (ref: Female)", "Some college (ref: High school or less)", "Associate's degree or more (ref: High school or less", "Separated (ref: Married)", "Single (ref: Married)", "No public benefits (ref: Receives public benefits)", "Unsure if receives public benefits (ref: Receives public benefits)", "Health insured (ref: Uninsured)", "Other employment (ref: Employed)", "Retired (ref: Employed)", "Unemployed (ref: Employed)", "Years lived in Flint", "Rented home (ref: Owned home)", "Unstable housing (ref: Owned home)", "Other housing (ref: Owned home)", "Other water source (ref: Flint tap water)", "Has vehicle (ref: No vehicle)", "Network size", "Adult uses filtered tap water (ref: Adult uses unfiltered water)", "Adult uses bottled water (ref: Adult uses unfiltered water)"), "Estimate" = summary(pool(child_bottled))$estimate[-1], "SE" = summary(pool(child_bottled))$std.error[-1], "Pval" = summary(pool(child_bottled))$p.value[-1])

#effect_estims <- rbind(effect_estims, subdat)
```

Since we fit quite a few models, we use the FDR to adjust for our multiple testing.
```{r, echo=FALSE, message=FALSE, warning=FALSE, eval = TRUE}
effect_estims <- effect_estims[-1,]
effect_estims$Pval[is.na(effect_estims$Pval)] <- 2*pnorm(abs(effect_estims$Estimate[is.na(effect_estims$Pval)]/effect_estims$SE[is.na(effect_estims$Pval)]), lower.tail = FALSE)
effect_estims$Lower <- effect_estims$Estimate - 1.96*effect_estims$SE
effect_estims$Upper <- effect_estims$Estimate + 1.96*effect_estims$SE

pretty_effects <- data.frame("Outcome" = effect_estims$Outcome, "Predictor" = effect_estims$Predictor, "Estimate" = exp(effect_estims$Estimate), "Lower" = exp(effect_estims$Lower), "Upper" = exp(effect_estims$Upper), "BHpvalue" = round(p.adjust(effect_estims$Pval, method="BH"), digits = 2))

pretty_effects$Label <- paste0(round(pretty_effects$Estimate, digits = 1), " (", round(pretty_effects$Lower, digits = 1), ", ", round(pretty_effects$Upper, digits = 1), ")")

pretty_effects2 <- dplyr::select(pretty_effects, Outcome, Predictor, Label, BHpvalue)

pretty_effects2[pretty_effects2$BHpvalue <= 0.05,]
```

```